---
title: " "
output:
  pdf_document:
    number_sections: TRUE
  html_document:
    code_folding: hide
documentclass: scrreprt
classoption: [parskip, oneside] # für drucken twoside
header-includes:
  - \usepackage[greek, ngerman, main=english]{babel}
  - \usepackage[utf8]{inputenc}
  - \usepackage[T1]{fontenc}
  - \usepackage{lmodern}
  - \usepackage[onehalfspacing]{setspace}
  - \usepackage[left=2.50cm, right=2.50cm, top=2.50cm, bottom=2.50cm, bindingoffset=10mm, includehead, includefoot]{geometry}
  - \usepackage[headsepline]{scrlayer-scrpage}
  - \usepackage{url}
  - \usepackage[backend=biber, style=authoryear, giveninits=true, maxbibnames=99, uniquename=init, maxcitenames=2, hyperref=true, date=year]{biblatex}
  #- \usepackage{bib/cellstyle_it}
  - \usepackage{xpatch}
  - \usepackage{csquotes}
  - \usepackage{amsmath}
  - \usepackage{listings}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{multirow}
  - \usepackage{rotating}
  - \usepackage{subfigure}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{acronym}
  - \usepackage{lipsum}
  - \usepackage{scrhack}
  - \usepackage{blindtext} %This package generates automatic text
  - \usepackage{epigraph} 
  - \emergencystretch=50pt
  - \clubpenalty = 10000
  - \widowpenalty = 10000
  - \displaywidowpenalty = 10000
  - \automark[section]{chapter}
  - \renewcommand*{\chaptermarkformat}{}
  - \renewcommand*{\sectionmarkformat}{}
  - \setkomafont{title}{\sffamily}
  - \setkomafont{disposition}{\usekomafont{title}}
  - \setkomafont{author}{\usekomafont{title}}
  - \setkomafont{date}{\usekomafont{title}}
  - \setkomafont{caption}{\sffamily\small}
  - \setkomafont{captionlabel}{\usekomafont{caption}\bfseries\small}
  - \setkomafont{pagehead}{\normalfont\scshape}
csl: bib/molecular-biology-of-the-cell.csl
#bibliography: [bib/packages.bib, bib/test.bib]
---
<!-- Based on Nicolas Peschkes LaTeX Template for MoBis, if you want further information on the packages just check it here: https://gitlab.com/npeschke/LaTeX-Template-IPMB -->

<!-- your title page -->

\begin{titlepage}
\centering
	{\Large Ruprecht-Karls-Universität Heidelberg\\
		Fakultät für Biowissenschaften\\
		Bachelorstudiengang Molekulare Biotechnologie\\}

	{\vspace{\stretch{2}}}
	{\usekomafont{title}

	{\Huge Data Analysis Report}

		{\Huge Topic 2, Group 1}

		{\Huge Supervisor: Carl Hermann}
		{\Huge Tutor: Wanjun Hu}

	}

	\vspace{\stretch{2}}
	{\Large Data Science Project SoSe 2022}

	\vspace{\stretch{2}}

	{\Large
		\begin{tabular}{rl}
			Autoreb & Ekin Ören, Yoana Onishtenko, Linh Trinh, Junona Sachov\\
			
			Abgabetermin & 20.07.2022\\
		\end{tabular}
	}

	\vspace{\stretch{1}}

\end{titlepage}


<!-- add your chapters here -->
\tableofcontents

```{r child = "/Users/lunona/Desktop/BRCA/GitHub/2022-topic-02-team-01/2022-topic-02-team-01/Report/chapter/abstract.Rmd"}
```

```{r child = "/Users/lunona/Desktop/BRCA/GitHub/2022-topic-02-team-01/2022-topic-02-team-01/Report/chapter/introduction.Rmd"}
```

```{r child = "/Users/lunona/Desktop/BRCA/GitHub/2022-topic-02-team-01/2022-topic-02-team-01/Report/chapter/material_and_methods.Rmd"}
```

```{r child = "/Users/lunona/Desktop/BRCA/GitHub/2022-topic-02-team-01/2022-topic-02-team-01/Report/chapter/results.Rmd"}
```

```{r child = "/Users/lunona/Desktop/BRCA/GitHub/2022-topic-02-team-01/2022-topic-02-team-01/Report/chapter/discussion.Rmd"}
```

# References

<div id="refs"></div>


```{r child = "/Users/lunona/Desktop/BRCA/GitHub/2022-topic-02-team-01/2022-topic-02-team-01/Report/chapter/appendix.Rmd"}
```