---
title: "BucketMenuChickenWings"
author: "Ekin Ören"
date: '2022-07-11'
output: html_document
---



# here we check if the brca in the SkinnyBoy and the SkinnyTT's the same people

What we learned: Sample IDS from skinnySADtt are in SB, but the ones from skinnyHAPPYtt are not

```{r papaya melon avocado cat, echo=FALSE}

SkinnyTT_Samplenames = colnames(pw_activity_SkinnyBRCA)

SkinnyBoy_Samplenames = colnames(SkinnyBoy)

SkinnySadTT_Samplenames = colnames(SkinnySadTT)

match(SkinnyTT_Samplenames, SkinnyBoy_Samplenames)

match(SkinnySadTT_Samplenames, SkinnyBoy_Samplenames)

```


Making volco plottes for tum and norm

```{r papaya avocado cat, echo=FALSE}




BRCApathways = pw_activity_SkinnyBRCA

tummyBRCApathways = BRCApathways[,1:111]
normieBRCApathways = BRCApathways[,111:222]


pee_BRCA_TUMvsNORM_pathways = c()
 for (i in 1:nrow(tummyBRCApathways)){
  pee_BRCA_TUMvsNORM_pathways[i] = wilcox.test(unlist(tummyBRCApathways[i,]),unlist(normieBRCApathways[i,]),paired = FALSE )$p.value
 }
p_BRCA_TUMvsNORM_pathways = pee_BRCA_TUMvsNORM_pathways*length(tummyBRCApathways[,1])

# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot
df_BRCA_TUMvsNORM_pathways = data.frame(rownames(normieBRCApathways), p_BRCA_TUMvsNORM_pathways)



foldchange_tum_norm = apply(tummyBRCApathways, 1, mean) - apply(normieBRCApathways, 1, mean)




#plotting time
df_volcanoplotBRCA_TUMvsNORM_pathways = data.frame (foldchange_tum_norm, p_BRCA_TUMvsNORM_pathways)
df_volcanoplotBRCA_TUMvsNORM_pathways$different = "NO"

df_volcanoplotBRCA_TUMvsNORM_pathways$different[df_volcanoplotBRCA_TUMvsNORM_pathways$foldchange_tum_norm > 0  &     df_volcanoplotBRCA_TUMvsNORM_pathways$p_BRCA_TUMvsNORM_pathways < 1.009123e-34]= "UP"

df_volcanoplotBRCA_TUMvsNORM_pathways$different[df_volcanoplotBRCA_TUMvsNORM_pathways$foldchange_tum_norm < 0 &      df_volcanoplotBRCA_TUMvsNORM_pathways$p_BRCA_TUMvsNORM_pathways < 1.009123e-34]= "DOWN"

df_volcanoplotBRCA_TUMvsNORM_pathways$different[which (row.names (df_volcanoplotBRCA_TUMvsNORM_pathways) == "YAMAZAKI_TCEB3_TARGETS_UP")]= "YAMAZAKI_TCEB3_TARGETS_UP"

Name = row.names(df_volcanoplotBRCA_TUMvsNORM_pathways)

#we plot 
ggplot (data = df_volcanoplotBRCA_TUMvsNORM_pathways, aes (x=foldchange_tum_norm, y=-log10(p_BRCA_TUMvsNORM_pathways), col = different)) + ggtitle("volcanoplot BRCA tumour vs norm pathways") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue", "red"))+ theme (text=element_text(size = 8) )

#tryyna put pathway names in this shit
  #geom_point() + geom_text(aes(label=ifelse(-log10(p_BRCA_TUMvsNORM_pathways)>30,as.character(Name),'')),hjust=0,vjust=0,)

JerryTheDataset = df_volcanoplotBRCA_TUMvsNORM_pathways[df_volcanoplotBRCA_TUMvsNORM_pathways$different == "UP" | df_volcanoplotBRCA_TUMvsNORM_pathways$different == "DOWN", ]

View(JerryTheDataset)

View(df_volcanoplotBRCA_TUMvsNORM_pathways)

rownames(JerryTheDataset)

saveRDS(JerryTheDataset,"C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/ DifferingPathwaysin_tumnormBRCA.rds")


 ggplot(nba, aes(x= MIN, y= PTS, colour="green", label=Name))+
  geom_point() +
  geom_text(aes(label=ifelse(PTS>24,as.character(Name),'')),hjust=0,vjust=0)


#saveRDS(df_volcanoplotBRCA_TUMvsNORM_pathways,"C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/ volcanoplotBRCA_TUMvsNORM_pathways.RDS")

```


# here we run the volcano plots for tumour comprisons for pathways (BRCA KIRC)

```{r papaya avocado cat, echo=FALSE}

#Loading and slicing the tumory data for further use
df_BRCA_pvalue = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/df_BRCA_pvalue.RDS")
VliveVdieVlivagen5 = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/Five_Tumory_Bois.RDS")
View(VliveVdieVlivagen5)
View(df_BRCA_pvalue)

BRCAtum = VliveVdieVlivagen5$BRCA$tumor
BRCAnorm = VliveVdieVlivagen5$BRCA$normal

KIRCtum = VliveVdieVlivagen5$KIRC$tumor
KIRCnorm = VliveVdieVlivagen5$KIRC$normal

LUADtum = VliveVdieVlivagen5$LUAD$tumor
LUADnorm = VliveVdieVlivagen5$LUAD$normal

PRADtum = VliveVdieVlivagen5$PRAD$tumor
PRADnorm = VliveVdieVlivagen5$PRAD$normal

THCAtum = VliveVdieVlivagen5$THCA$tumor
THCAnorm = VliveVdieVlivagen5$THCA$normal

```


```{r melon avocado, echo=FALSE}

#first we cbind the tum and norm of this tumourtype and clean the names

KIRCtum_norm_gebunden = cbind(KIRCtum, KIRCnorm)

gene_identifier <- strsplit(rownames(KIRCtum_norm_gebunden), split = "|", fixed = TRUE)

genes_ens<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[1]
  return(ens)
})

genes_symbol<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[2]
  return(ens)
})

genes_ens_vn <- strsplit(genes_ens, split = ".", fixed = TRUE)
genes_ensembl <- sapply(genes_ens_vn, FUN = function(genes_ens_vn){
  ens = genes_ens_vn[1]
  return(ens)
})

rownames(KIRCtum_norm_gebunden) = genes_ensembl




#now we do GSVA to get pathway scores for this tumour type

library(GSVA)
library(BiocParallel)
library(msigdbr)
library(ggplot2)
library(uwot)

params <- MulticoreParam(workers = multicoreWorkers())

pw_activity <- gsva(expr = as.matrix(KIRCtum_norm_gebunden),
                    gset.idx.list = AllPathwaysEns_List,
                    kcdf = "Gaussian",
                    min.sz = 15,
                    BPPARAM = params)

KIRCpathways = pw_activity

KIRCpathways = KIRCpathways[match(rownames(BRCApathways), rownames(KIRCpathways)),]



#now we compute the p scores of the GSVA scores

BRCApathways = pw_activity_SkinnyBRCA


pee_BRCAvsKIRCpathways = c()
 for (i in 1:nrow(BRCApathways)){
  pee_BRCAvsKIRCpathways[i] = wilcox.test(unlist(BRCApathways[i,]),unlist(KIRCpathways[i,]),paired = FALSE )$p.value
 }
p_BRCAvsKIRCpathways = pee_BRCAvsKIRCpathways*length(BRCApathways[,1])


# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot
df_pBRCAvsKIRCpathways = data.frame(rownames(KIRCpathways), p_BRCAvsKIRCpathways)

View(df_pBRCAvsKIRCpathways)


```



```{r papaya avocado cat, echo=FALSE}

foldchange11 = apply(BRCApathways, 1, mean) - apply(KIRCpathways, 1, mean)

```



```{r papaya avocado cat, echo=FALSE}


df_volcanoplotBRCAvsKIRCpathways = data.frame (foldchange11, p_BRCAvsKIRCpathways)
df_volcanoplotBRCAvsKIRCpathways$different = "NO"

df_volcanoplotBRCAvsKIRCpathways$different[df_volcanoplotBRCAvsKIRCpathways$foldchange1 > 0  &     df_volcanoplotBRCAvsKIRCpathways$p_BRCAvsKIRCpathways < 170]= "UP"

df_volcanoplotBRCAvsKIRCpathways$different[df_volcanoplotBRCAvsKIRCpathways$foldchange1 < 0 &      df_volcanoplotBRCAvsKIRCpathways$p_BRCAvsKIRCpathways < 170]= "DOWN"

df_volcanoplotBRCAvsKIRCpathways$different[which (row.names (df_volcanoplotBRCAvsKIRCpathways) == "YAMAZAKI_TCEB3_TARGETS_UP")]= "YAMAZAKI_TCEB3_TARGETS_UP"

#we plot 
ggplot (data = df_volcanoplotBRCAvsKIRCpathways, aes (x=foldchange11, y=-log10(p_BRCAvsKIRCpathways), col = different)) + ggtitle("volcanoplot BRCA tumour vs KIRC tumour pathways") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue", "red"))+ theme (text=element_text(size = 8) )

View(HarryTheDataset)

HarryTheDataset = df_volcanoplotBRCAvsKIRCpathways[df_volcanoplotBRCAvsKIRCpathways$different == "UP" | df_volcanoplotBRCAvsKIRCpathways$different == "DOWN", ]

match(rownames(JerryTheDataset), rownames(HarryTheDataset))

!is.na(match(rownames(JerryTheDataset), rownames(HarryTheDataset))) # gives 52 and 55



DifferingPathwaysin_tumnorm_BRCAKIRC = HarryTheDataset[c(52,55),]


```


BRCA LUAD

```{r melon avocado, echo=FALSE}

#first we cbind the tum and norm of this tumourtype and clean the names

LUADtum_norm_gebunden = cbind(LUADtum, LUADnorm)

gene_identifier <- strsplit(rownames(LUADtum_norm_gebunden), split = "|", fixed = TRUE)

genes_ens<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[1]
  return(ens)
})

genes_symbol<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[2]
  return(ens)
})

genes_ens_vn <- strsplit(genes_ens, split = ".", fixed = TRUE)
genes_ensembl <- sapply(genes_ens_vn, FUN = function(genes_ens_vn){
  ens = genes_ens_vn[1]
  return(ens)
})

rownames(LUADtum_norm_gebunden) = genes_ensembl

View(AllPathwaysEns_List)

#now we do GSVA to get pathway scores for this tumour type

library(GSVA)
library(BiocParallel)
library(msigdbr)
library(ggplot2)
library(uwot)

params <- MulticoreParam(workers = multicoreWorkers())

pw_activity <- gsva(expr = as.matrix(LUADtum_norm_gebunden),
                    gset.idx.list = AllPathwaysEns_List,
                    kcdf = "Gaussian",
                    min.sz = 15,
                    BPPARAM = params)

LUADpathways = pw_activity

LUADpathways = LUADpathways[match(rownames(BRCApathways), rownames(LUADpathways)),]

View(LUADpathways)

#now we compute the p scores of the GSVA scores



BRCApathways = pw_activity_SkinnyBRCA


pee_BRCAvsLUADpathways = c()
 for (i in 1:nrow(BRCApathways)){
  pee_BRCAvsLUADpathways[i] = wilcox.test(unlist(BRCApathways[i,]),unlist(LUADpathways[i,]),paired = FALSE )$p.value
 }
p_BRCAvsLUADpathways = pee_BRCAvsLUADpathways*length(BRCApathways[,1])



# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot
df_pBRCAvsLUADpathways = data.frame(rownames(LUADpathways), p_BRCAvsLUADpathways)

View(df_pBRCAvsLUADpathways)

foldchange22 = apply(BRCApathways, 1, mean) - apply(LUADpathways, 1, mean)



#plot time

df_volcanoplotBRCAvsLUADpathways = data.frame (foldchange22, p_BRCAvsLUADpathways)
df_volcanoplotBRCAvsLUADpathways$different = "NO"

df_volcanoplotBRCAvsLUADpathways$different[df_volcanoplotBRCAvsLUADpathways$foldchange1 > 0  &     df_volcanoplotBRCAvsLUADpathways$p_BRCAvsLUADpathways < 170]= "UP"

df_volcanoplotBRCAvsLUADpathways$different[df_volcanoplotBRCAvsLUADpathways$foldchange1 < 0 &      df_volcanoplotBRCAvsLUADpathways$p_BRCAvsLUADpathways < 170]= "DOWN"

#we plot 
ggplot (data = df_volcanoplotBRCAvsLUADpathways, aes (x=foldchange22, y=-log10(p_BRCAvsLUADpathways), col = different)) + ggtitle("volcanoplot BRCA tumour vs KIRC tumour pathways") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )

View(df_volcanoplotBRCAvsLUADpathways)


```



BRCA VS PRAD

```{r melon avocado, echo=FALSE}

#first we cbind the tum and norm of this tumourtype and clean the names

PRADtum_norm_gebunden = cbind(PRADtum, PRADnorm)

gene_identifier <- strsplit(rownames(PRADtum_norm_gebunden), split = "|", fixed = TRUE)

genes_ens<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[1]
  return(ens)
})

genes_symbol<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[2]
  return(ens)
})

genes_ens_vn <- strsplit(genes_ens, split = ".", fixed = TRUE)
genes_ensembl <- sapply(genes_ens_vn, FUN = function(genes_ens_vn){
  ens = genes_ens_vn[1]
  return(ens)
})

rownames(PRADtum_norm_gebunden) = genes_ensembl



#now we do GSVA to get pathway scores for this tumour type

library(GSVA)
library(BiocParallel)
library(msigdbr)
library(ggplot2)
library(uwot)

params <- MulticoreParam(workers = multicoreWorkers())

pw_activity <- gsva(expr = as.matrix(PRADtum_norm_gebunden),
                    gset.idx.list = AllPathwaysEns_List,
                    kcdf = "Gaussian",
                    min.sz = 15,
                    BPPARAM = params)

PRADpathways = pw_activity

PRADpathways = PRADpathways[match(rownames(BRCApathways), rownames(PRADpathways)),]

View(PRADpathways)

#now we compute the p scores of the GSVA scores



BRCApathways = pw_activity_SkinnyBRCA


pee_BRCAvsPRADpathways = c()
 for (i in 1:nrow(BRCApathways)){
  pee_BRCAvsPRADpathways[i] = wilcox.test(unlist(BRCApathways[i,]),unlist(PRADpathways[i,]),paired = FALSE )$p.value
 }
p_BRCAvsPRADpathways = pee_BRCAvsPRADpathways*length(BRCApathways[,1])


# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot
df_pBRCAvsPRADpathways = data.frame(rownames(PRADpathways), p_BRCAvsPRADpathways)

View(df_pBRCAvsPRADpathways)

foldchange33 = apply(BRCApathways, 1, mean) - apply(PRADpathways, 1, mean)



#plot time

df_volcanoplotBRCAvsPRADpathways = data.frame (foldchange33, p_BRCAvsPRADpathways)
df_volcanoplotBRCAvsPRADpathways$different = "NO"

df_volcanoplotBRCAvsPRADpathways$different[df_volcanoplotBRCAvsPRADpathways$foldchange33 > 0  &     df_volcanoplotBRCAvsPRADpathways$p_BRCAvsPRADpathways < 300]= "UP"

df_volcanoplotBRCAvsPRADpathways$different[df_volcanoplotBRCAvsPRADpathways$foldchange33 < 0 &      df_volcanoplotBRCAvsPRADpathways$p_BRCAvsPRADpathways < 300]= "DOWN"


df_volcanoplotBRCAvsPRADpathways$different[which (row.names (df_volcanoplotBRCAvsPRADpathways) == "YAMAZAKI_TCEB3_TARGETS_UP")]= "YAMAZAKI_TCEB3_TARGETS_UP"



#we plot 
ggplot (data = df_volcanoplotBRCAvsPRADpathways, aes (x=foldchange33, y=-log10(p_BRCAvsPRADpathways), col = different)) + ggtitle("volcanoplot BRCA tumour vs PRAD tumour pathways") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue", "red"))+ theme (text=element_text(size = 8) )

View(df_volcanoplotBRCAvsPRADpathways)

MerryTheDataset = df_volcanoplotBRCAvsPRADpathways[df_volcanoplotBRCAvsPRADpathways$different == "UP" | df_volcanoplotBRCAvsPRADpathways$different == "DOWN", ]

match(rownames(JerryTheDataset), rownames(MerryTheDataset))  # gives 

bue = unique(match(rownames(JerryTheDataset), rownames(MerryTheDataset)) )

bue1 = bue [-1]

DifferingPathwaysin_tumnorm_BRCAPRAD = MerryTheDataset[bue1,]

View(DifferingPathwaysin_tumnorm_BRCAPRAD)

```



BRCA VS THCA

```{r melon avocado, echo=FALSE}

#first we cbind the tum and norm of this tumourtype and clean the names

THCAtum_norm_gebunden = cbind(THCAtum, THCAnorm)

gene_identifier <- strsplit(rownames(THCAtum_norm_gebunden), split = "|", fixed = TRUE)

genes_ens<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[1]
  return(ens)
})

genes_symbol<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[2]
  return(ens)
})

genes_ens_vn <- strsplit(genes_ens, split = ".", fixed = TRUE)
genes_ensembl <- sapply(genes_ens_vn, FUN = function(genes_ens_vn){
  ens = genes_ens_vn[1]
  return(ens)
})

rownames(THCAtum_norm_gebunden) = genes_ensembl



#now we do GSVA to get pathway scores for this tumour type

library(GSVA)
library(BiocParallel)
library(msigdbr)
library(ggplot2)
library(uwot)

params <- MulticoreParam(workers = multicoreWorkers())

pw_activity <- gsva(expr = as.matrix(THCAtum_norm_gebunden),
                    gset.idx.list = AllPathwaysEns_List,
                    kcdf = "Gaussian",
                    min.sz = 15,
                    BPPARAM = params)

THCApathways = pw_activity

BRCApathways = pw_activity_SkinnyBRCA[,1:111]

THCApathways = THCApathways[match(rownames(BRCApathways), rownames(THCApathways)),]

#now we compute the p scores of the GSVA scores



#try1
#pee_BRCAvsTHCApathways = c()
 #for (i in 1:nrow(as.numeric(nrow(BRCApathways)))){
  #pee_BRCAvsTHCApathways[i] = wilcox.test(as.numeric(unlist(THCApathways[i,])),as.numeric(unlist(BRCApathways[i,])),paired = FALSE )$p.value
 #}
#p_BRCAvsTHCApathways = pee_BRCAvsTHCApathways*length(BRCApathways[,1])


#try2
pee_BRCAvsTHCApathways = c()
 for (i in 1:nrow(BRCApathways)){
  pee_BRCAvsTHCApathways[i] = wilcox.test(as.numeric(THCApathways[i,]),as.numeric(BRCApathways[i,]),paired = TRUE )$p.value
 }
p_BRCAvsTHCApathways = pee_BRCAvsTHCApathways*length(BRCApathways[,1])

saveRDS(BRCApathways, "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/BRCApathwaysWilcox.rds")
saveRDS(THCApathways, "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/THCApathwaysWilcox.rds")

#Original

pee_BRCAvsTHCApathways = c()
 for (i in 1:nrow(BRCApathways)){
  pee_BRCAvsTHCApathways[i] = wilcox.test(unlist(THCApathways[i,]),unlist(BRCApathways[i,]),paired = FALSE )$p.value
 }
p_BRCAvsTHCApathways = pee_BRCAvsTHCApathways*length(BRCApathways[,1])


# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot
df_pBRCAvsTHCApathways = data.frame(rownames(THCApathways), p_BRCAvsTHCApathways) 

View(df_pBRCAvsTHCApathways)

foldchange44 = apply(BRCApathways, 1, mean) - apply(THCApathways, 1, mean)



#plot time

df_volcanoplotBRCAvsTHCApathways = data.frame (foldchange44, p_BRCAvsTHCApathways)
df_volcanoplotBRCAvsTHCApathways$different = "NO"

df_volcanoplotBRCAvsTHCApathways$different[df_volcanoplotBRCAvsTHCApathways$foldchange44 > 0  &     df_volcanoplotBRCAvsTHCApathways$p_BRCAvsTHCApathways < 300]= "UP"

df_volcanoplotBRCAvsTHCApathways$different[df_volcanoplotBRCAvsTHCApathways$foldchange44 < 0 &      df_volcanoplotBRCAvsTHCApathways$p_BRCAvsTHCApathways < 300]= "DOWN"

#we plot 
ggplot (data = df_volcanoplotBRCAvsTHCApathways, aes (x=foldchange44, y=-log10(p_BRCAvsTHCApathways), col = different)) + ggtitle("volcanoplot BRCA tumour vs PRAD tumour pathways") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )

View(df_volcanoplotBRCAvsTHCApathways)


```



TRY at comparison of BRCA tum and all skinnyboy tums

```{r melon avocado, echo=FALSE}


Aliali = match (colnames(pw_activity_SkinnyBRCA) ,  colnames(PathwayMatrixDeluxePremium))
  
Veliveli = Aliali[which(!is.na(match (colnames(pw_activity_SkinnyBRCA) ,  colnames(PathwayMatrixDeluxePremium))))]

BRCApathways1 = PathwayMatrixDeluxePremium[,Veliveli]



MUTUALpathways = PathwayMatrixDeluxePremium[match(rownames(BRCApathways1), rownames(PathwayMatrixDeluxePremium)),]

View(MUTUALpathways)




#now we compute the p scores of the GSVA scores

pee_BRCAvsSBpathways = c()
 for (i in 1:nrow(BRCApathways1)){
  pee_BRCAvsSBpathways[i] = wilcox.test(unlist(BRCApathways1[i,]),unlist(MUTUALpathways[i,]),paired = FALSE )$p.value
 }
p_BRCAvsSBpathways = pee_BRCAvsSBpathways*length(BRCApathways1[,1])


# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot
df_pBRCAvsSBpathways = data.frame(rownames(MUTUALpathways), p_BRCAvsSBpathways) 

foldchange666 = apply(BRCApathways1, 1, mean) - apply(MUTUALpathways, 1, mean)



#plot time

df_volcanoplotBRCAvsSBpathways = data.frame (foldchange666, p_BRCAvsSBpathways)
df_volcanoplotBRCAvsSBpathways$different = "NO"

df_volcanoplotBRCAvsSBpathways$different[df_volcanoplotBRCAvsSBpathways$foldchange666 > 0  &     df_volcanoplotBRCAvsSBpathways$p_BRCAvsSBpathways < 0.009]= "UP"

df_volcanoplotBRCAvsSBpathways$different[df_volcanoplotBRCAvsSBpathways$foldchange666 < 0 &      df_volcanoplotBRCAvsSBpathways$p_BRCAvsSBpathways < 0.009]= "DOWN"

#df_volcanoplotBRCAvsSBpathways$different[which (row.names (df_volcanoplotBRCAvsSBpathways) == "YAMAZAKI_TCEB3_TARGETS_UP")]= "YAMAZAKI_TCEB3_TARGETS_UP"

#we plot 
library(ggplot2)

which (row.names (df_volcanoplotBRCAvsSBpathways) == "YAMAZAKI_TCEB3_TARGETS_UP")


ggplot (data = df_volcanoplotBRCAvsSBpathways, aes (x=foldchange666, y=-log10(p_BRCAvsSBpathways), col = different)) + ggtitle("volcanoplot BRCA tumour vs SB pathways") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue", "red"))+ theme (text=element_text(size = 8))


View(df_volcanoplotBRCAvsSBpathways)



```






