---
title: "VolcanoPlotter"
author: "Ekin Ören"
date: '2022-06-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Volcano plot is a type of gg plot to be able to use it you need to call a library. 

```{r pressure, echo=FALSE}

#Using ggplot
install.packages("Rcmdr", dependencies = TRUE)
install.packages("tidyverse")
library(ggplot2)

```


p <- ggplot(data=de, aes(x=log2FoldChange, y=-log10(pvalue))) + geom_point() + theme_minimal()

```{r my money dont jiggle jiggle it folds, include=FALSE}



p <- ggplot(data=MetabolicPathwayDF[,1], aes(x=LoggyFoldChange, y=-log10(MetabolicPathwayDF[,1]))) + geom_point() + theme_minimal()

```




```{r my money dont jiggle jiggle it folds, include=FALSE}

library(gtools)

foldchange(A,B)
Log2FoldChange = log2(foldchange(A,B))




A1  = A + 9.9658 + 0.01

B1 = B + 9.9658 + 0.01



foldchange(A,B)[1:10]
foldchange(A1,B1)[1:10]

log2(foldchange(A,B))
log2(foldchange(A1,B1))

absABfold = abs(foldchange(A,B)[1:10])

log2(absABfold)




```



Calculate fold change between mean of gene exp of one gene
```{r my money dont jiggle jiggle it folds, include=FALSE}

A = SkinnyBoy[,1]
B = SkinnyBoy[,2]
RawFoldChange = c()


for(i in 1:length(SkinnyBoy[,1])){
  
  RawFoldChange[i] =  (B[i] - A[i])/A[i]
  
}

#Not sure if these are correct
LoggyFoldChange = log2(RawFoldChange)
LoggyFoldChangeClean = LoggyFoldChange[!is.nan(LoggyFoldChange)]



```
THEORY BEHİND LOG CHANGE:

For quantities A and B, the fold change is given as (B − A)/A, or equivalently B/A − 1. This formulation has appealing properties such as no change being equal to zero, a 100% increase is equal to 1, and a 100% decrease is equal to −1.

Let's say there are 50 read counts in control and 100 read counts in treatment for gene A. This means gene A is expressing twice in treatment as compared to control (100 divided by 50 =2) or fold change is 2. This works well for over expressed genes as the number directly corresponds to how many times a gene is over-expressed. But when it is other way round (i.e, treatment 50, control 100), the value of fold change will be 0.5 (all under expressed genes will have values between 0 to 1, while over expressed genes will have values from 1 to infinity). To make this leveled, we use log2 for expressing the fold change. I.e, log2 of 2 is 1 and log2 of 0.5 is -1.




```{r my money dont jiggle jiggle it folds, include=FALSE}

# Download the data from github (click the "raw" button, save as a text file called "results.txt").
# https://gist.github.com/stephenturner/806e31fce55a8b7175af
res <- read.table("results.txt", header=TRUE)
head(res)

# Make a basic volcano plot
with(res, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-2.5,2)))

# Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(res, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
with(subset(res, abs(log2FoldChange)>1), points(log2FoldChange, -log10(pvalue), pch=20, col="orange"))
with(subset(res, padj<.05 & abs(log2FoldChange)>1), points(log2FoldChange, -log10(pvalue), pch=20, col="green"))

# Label points with the textxy function from the calibrate plot
library(calibrate)
with(subset(res, padj<.05 & abs(log2FoldChange)>1), textxy(log2FoldChange, -log10(pvalue), labs=Gene, cex=.8))



```


































```{r my money dont jiggle jiggle it folds, include=FALSE}



```







