---
title: "Pathway_Volcano"
author: "Ekin Ören"
date: '2022-06-19'
output: html_document
---

#Loading and slicing the tumory data for further use
```{r Load DF's, }

#Loading and slicing the tumory data for further use
df_BRCA_pvalue = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/df_BRCA_pvalue.RDS")
VliveVdieVlivagen5 = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/Five_Tumory_Bois.RDS")
View(VliveVdieVlivagen5)
View(df_BRCA_pvalue)

BRCAtum = VliveVdieVlivagen5$BRCA$tumor
BRCAnorm = VliveVdieVlivagen5$BRCA$normal

KIRCtum = VliveVdieVlivagen5$KIRC$tumor
KIRCnorm = VliveVdieVlivagen5$KIRC$normal

LUADtum = VliveVdieVlivagen5$LUAD$tumor
LUADnorm = VliveVdieVlivagen5$LUAD$normal

PRADtum = VliveVdieVlivagen5$PRAD$tumor
PRADnorm = VliveVdieVlivagen5$PRAD$normal

THCAtum = VliveVdieVlivagen5$THCA$tumor
THCAnorm = VliveVdieVlivagen5$THCA$normal
```



#Here we compute the p values of all cancer types with comparison to BRCA
```{r compute p-values for each genes}

pee_BRCAvsKIRC = c()
 for (i in 1:nrow(BRCAtum)){
  pee_BRCAvsKIRC[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(KIRCtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsKIRC = pee_BRCAvsKIRC*length(BRCAtum[,1])


pee_BRCAvsLUAD = c()
 for (i in 1:nrow(BRCAtum)){
  pee_BRCAvsLUAD[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(LUADtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsLUAD = pee_BRCAvsLUAD*length(BRCAtum[,1])


pee_BRCAvsPRAD = c()
 for (i in 1:nrow(BRCAtum)){
  pee_BRCAvsPRAD[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(PRADtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsPRAD = pee_BRCAvsPRAD*length(BRCAtum[,1])


pee_BRCAvsTHCA = c()
 for (i in 1:nrow(BRCAtum)){
  pee_BRCAvsTHCA[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(THCAtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsTHCA = pee_BRCAvsTHCA*length(BRCAtum[,1])

# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot

df_pBRCAvsKIRC = data.frame(rownames(BRCAtum), p_BRCAvsKIRC)
df_pBRCAvsLUAD = data.frame(rownames(BRCAtum), p_BRCAvsLUAD)
df_pBRCAvsPRAD = data.frame(rownames(BRCAtum), p_BRCAvsPRAD)
df_pBRCAvsTHCA = data.frame(rownames(BRCAtum), p_BRCAvsTHCA)

```



#here we find the highest and lowest P values in previous comparisons and draw out the gene
#we don't need to calculate additional correlation value because we already computed p value
#we must however also consider fold change. so we will make volcano plots for each of these comparisons
#to do this we first need to compute fold change

```{r Foldchange, }

#each row (row not column ı checked)
foldchange1 = apply(BRCAtum, 1, mean) - apply(KIRCtum, 1, mean)
log2foldchange1 = log2(foldchange1)

foldchange2 = apply(BRCAtum, 1, mean) - apply(LUADtum, 1, mean)
log2foldchange2 = log2(foldchange2)

foldchange3 = apply(BRCAtum, 1, mean) - apply(PRADtum, 1, mean)
log2foldchange3 = log2(foldchange3)

foldchange4 = apply(BRCAtum, 1, mean) - apply(THCAtum, 1, mean)
log2foldchange4 = log2(foldchange4)

#PROBLEM: NaN's get produced here (PLS DEBUG)

```



#here we make the volcano plots for all cancer types
```{r Vocano yea, }


df_volcanoplotBRCAvsKIRC = data.frame (foldchange1, p_BRCAvsKIRC)
df_volcanoplotBRCAvsKIRC$different = "NO"

df_volcanoplotBRCAvsKIRC$different[df_volcanoplotBRCAvsKIRC$foldchange1 > 0.1  &     df_volcanoplotBRCAvsKIRC$p_BRCAvsKIRC < 0.00001]= "UP"

df_volcanoplotBRCAvsKIRC$different[df_volcanoplotBRCAvsKIRC$foldchange1 < 0.1 &      df_volcanoplotBRCAvsKIRC$p_BRCAvsKIRC < 0.00001]= "DOWN"

#we plot 
ggplot (data = df_volcanoplotBRCAvsKIRC, aes (x=foldchange1, y=-log10(p_BRCAvsKIRC), col = different)) + ggtitle("volcanoplot BRCA tumour vs KIRC tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )




df_volcanoplotBRCAvsLUAD = data.frame (foldchange2, p_BRCAvsLUAD)
df_volcanoplotBRCAvsLUAD$different = "NO"

df_volcanoplotBRCAvsLUAD$different[df_volcanoplotBRCAvsLUAD$foldchange2 > 0.1  &     df_volcanoplotBRCAvsLUAD$p_BRCAvsLUAD < 0.00001]= "UP"

df_volcanoplotBRCAvsLUAD$different[df_volcanoplotBRCAvsLUAD$foldchange2 < 0.1 &      df_volcanoplotBRCAvsLUAD$p_BRCAvsLUAD < 0.00001]= "DOWN"

#we plot
ggplot (data = df_volcanoplotBRCAvsLUAD, aes (x=foldchange2, y=-log10(p_BRCAvsLUAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs LUAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )




df_volcanoplotBRCAvsPRAD = data.frame (foldchange3, p_BRCAvsPRAD)
df_volcanoplotBRCAvsPRAD$different = "NO"

df_volcanoplotBRCAvsPRAD$different[df_volcanoplotBRCAvsPRAD$foldchange3 > 0.1  &     df_volcanoplotBRCAvsPRAD$p_BRCAvsPRAD < 0.00001]= "UP"

df_volcanoplotBRCAvsPRAD$different[df_volcanoplotBRCAvsPRAD$foldchange3 < 0.1 &      df_volcanoplotBRCAvsPRAD$p_BRCAvsPRAD < 0.00001]= "DOWN"

#we plot 
ggplot (data = df_volcanoplotBRCAvsPRAD, aes (x=foldchange3, y=-log10(p_BRCAvsPRAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs PRAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )




df_volcanoplotBRCAvsTHCA = data.frame (foldchange4, p_BRCAvsTHCA)
df_volcanoplotBRCAvsTHCA$different = "NO"

df_volcanoplotBRCAvsTHCA$different[df_volcanoplotBRCAvsTHCA$foldchange4 > 0.1  &     df_volcanoplotBRCAvsTHCA$p_BRCAvsTHCA < 0.00001]= "UP"

df_volcanoplotBRCAvsTHCA$different[df_volcanoplotBRCAvsTHCA$foldchange4 < 0.1 &      df_volcanoplotBRCAvsTHCA$p_BRCAvsTHCA < 0.00001]= "DOWN"

#we plot 
ggplot (data = df_volcanoplotBRCAvsTHCA, aes (x=foldchange4, y=-log10(p_BRCAvsTHCA), col = different)) + ggtitle("volcanoplot BRCA tumour vs THCA tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )


```

```{r Plots (with normal grenze), }


#Recap of all of the plots:

ggplot (data = df_volcanoplotBRCAvsKIRC, aes (x=foldchange1, y=-log10(p_BRCAvsKIRC), col = different)) + ggtitle("volcanoplot BRCA tumour vs KIRC tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )


ggplot (data = df_volcanoplotBRCAvsLUAD, aes (x=foldchange2, y=-log10(p_BRCAvsLUAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs LUAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )


ggplot (data = df_volcanoplotBRCAvsPRAD, aes (x=foldchange3, y=-log10(p_BRCAvsPRAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs PRAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )

ggplot (data = df_volcanoplotBRCAvsTHCA, aes (x=foldchange4, y=-log10(p_BRCAvsTHCA), col = different)) + ggtitle("volcanoplot BRCA tumour vs THCA tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )


```



```{r Plots (with normal grenze and with strict), }

#Stricc Plot

df_volcanoplotBRCAvsKIRCstrict = data.frame (foldchange1, p_BRCAvsKIRC)
df_volcanoplotBRCAvsKIRCstrict$different = "NO"

df_volcanoplotBRCAvsKIRCstrict$different[df_volcanoplotBRCAvsKIRCstrict$foldchange1 > 0.1  &     df_volcanoplotBRCAvsKIRCstrict$p_BRCAvsKIRC < 0.0000000000000000000001]= "UP"

df_volcanoplotBRCAvsKIRCstrict$different[df_volcanoplotBRCAvsKIRCstrict$foldchange1 < 0.1 &      df_volcanoplotBRCAvsKIRCstrict$p_BRCAvsKIRC < 0.00000000000000000000001]= "DOWN"

#we plot 
ggplot (data = df_volcanoplotBRCAvsKIRC, aes (x=foldchange1, y=-log10(p_BRCAvsKIRC), col = different)) + ggtitle("volcanoplot BRCA tumour vs KIRC tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )

ggplot (data = df_volcanoplotBRCAvsKIRCstrict, aes (x=foldchange1, y=-log10(p_BRCAvsKIRC), col = different)) + ggtitle("volcanoplot BRCA tumour vs KIRC tumour STRICT") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )

#As you can see in the "STRİCT" df we have only the most differentially expressed genes identified as up and down.
#we can make barplots with those later


df_volcanoplotBRCAvsLUADstrict = data.frame (foldchange2, p_BRCAvsLUAD)
df_volcanoplotBRCAvsLUADstrict$different = "NO"

df_volcanoplotBRCAvsLUADstrict$different[df_volcanoplotBRCAvsLUADstrict$foldchange2 > 0.1  &     df_volcanoplotBRCAvsLUADstrict$p_BRCAvsLUAD < 0.000000000000001]= "UP"

df_volcanoplotBRCAvsLUADstrict$different[df_volcanoplotBRCAvsLUADstrict$foldchange2 < 0.1 &      df_volcanoplotBRCAvsLUADstrict$p_BRCAvsLUAD <  0.0000000000000001]= "DOWN"


ggplot (data = df_volcanoplotBRCAvsLUAD, aes (x=foldchange2, y=-log10(p_BRCAvsLUAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs LUAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )

ggplot (data = df_volcanoplotBRCAvsLUADstrict, aes (x=foldchange2, y=-log10(p_BRCAvsLUAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs LUAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )





```




Here we explore the genes that were distinct in the plots we saw as an output of the chunk above

```{r get genes, }

Uppers1 <- df_volcanoplotBRCAvsKIRCstrict[ which(df_volcanoplotBRCAvsKIRCstrict$different =='UP'),]
#View(Uppers) #72 genes

Downers1 <- df_volcanoplotBRCAvsKIRCstrict[ which(df_volcanoplotBRCAvsKIRCstrict$different =='DOWN'),]
#View(Downers) #65 genes

UppersExpression1 =  SkinnyBoy[(match(rownames(Uppers1), rownames(SkinnyBoy))), ]
UppersExpression1 = na.omit(UppersExpression1)

DownersExpression1 =  SkinnyBoy[(match(rownames(Downers1), rownames(SkinnyBoy))), ]
DownersExpression1 = na.omit(DownersExpression1)

Uppers2 <- df_volcanoplotBRCAvsLUADstrict[ which(df_volcanoplotBRCAvsLUADstrict$different =='UP'),]
#View(Uppers) #72 genes

Downers2 <- df_volcanoplotBRCAvsLUADstrict[ which(df_volcanoplotBRCAvsLUADstrict$different =='DOWN'),]
#View(Downers) #65 genes

UppersExpression2 =  SkinnyBoy[(match(rownames(Uppers2), rownames(SkinnyBoy))), ]
UppersExpression2 = na.omit(UppersExpression2)

DownersExpression2 =  SkinnyBoy[(match(rownames(Downers2), rownames(SkinnyBoy))), ]
DownersExpression2 = na.omit(DownersExpression2)

```




In this chunk we look at the corrolation values. howeverwe ended up making a different corrolation analysis after an update to the pathway matrix
so this chunk is used only as a diagnostic tool

```{r lookat corrolation, }
#
UppersMeanExp1 = apply(UppersExpression1, 1, mean)
DownersMeanExp1 = apply(DownersExpression1, 1, mean)

UppersMeanExp2 = apply(UppersExpression2, 1, mean)
DownersMeanExp2 = apply(DownersExpression2, 1, mean)


Testpresso1 =  SkinnyBoy[(match(rownames(df_BRCA_pvalue), rownames(SkinnyBoy))), ]
Testpresso1 = na.omit(Testpresso1)


plot(UppersMeanExp1[1:100], Testpresso1[1:100], pch = 19, col = "lightblue")


View(df_BRCA_pvalue)





n = cor.test(UppersMeanExp1[1:100], UppersMeanExp1[1:100])
n$estimate

a = cor.test(UppersMeanExp1[1:100], UppersMeanExp2[1:100])
a$estimate

b = cor.test(UppersMeanExp1[1:100], DownersMeanExp1[1:100])
b$estimate


View(UppersExpression1)


#corrolation plots of first 200 upper genes in several patients
for (i in 1:10) {
  
  plot(UppersExpression1[1:200,i], UppersExpression1[1:200, i+1], pch = 19, col = "lightblue")

}

UpperGene_AllPatientsCor = c()

for (i in 1:(length(UppersExpression1[,1])-1)) {
  
  z = cor.test(UppersExpression1[,i], UppersExpression1[, i+1], pch = 19, col = "lightblue")
  UpperGene_AllPatientsCor[i] = z$estimate

}

hist(UpperGene_AllPatientsCor, breaks = 40)



```


```{r compute p-values for each genes}

#Here we compute the p values of all cancer types with comparison to BRCA

pea_BRCAvsKIRC = c()
 for (i in 1:nrow(BRCAtum)){
  pea_BRCAvsKIRC[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(KIRCtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsKIRC = pea_BRCAvsKIRC*length(BRCAtum[,1])


pea_BRCAvsLUAD = c()
 for (i in 1:nrow(BRCAtum)){
  pea_BRCAvsLUAD[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(LUADtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsLUAD = pea_BRCAvsLUAD*length(BRCAtum[,1])


pea_BRCAvsPRAD = c()
 for (i in 1:nrow(BRCAtum)){
  pea_BRCAvsPRAD[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(PRADtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsPRAD = pea_BRCAvsPRAD*length(BRCAtum[,1])


pea_BRCAvsTHCA = c()
 for (i in 1:nrow(BRCAtum)){
  pea_BRCAvsTHCA[i] = wilcox.test(unlist(BRCAtum[i,]),unlist(THCAtum[i,]),paired = FALSE )$p.value
 }
p_BRCAvsTHCA = pea_BRCAvsTHCA*length(BRCAtum[,1])

# in the end we create dataframes iwth gene names and p values for easy matching of the gene to the values in plot

df_pBRCAvsKIRC = data.frame(rownames(BRCAtum), p_BRCAvsKIRC)
df_pBRCAvsLUAD = data.frame(rownames(BRCAtum), p_BRCAvsLUAD)
df_pBRCAvsPRAD = data.frame(rownames(BRCAtum), p_BRCAvsPRAD)
df_pBRCAvsTHCA = data.frame(rownames(BRCAtum), p_BRCAvsTHCA)

```

```{r Foldchange, }

#here we find the highest and lowest P values in previous comparisons and draw out the gene
#we don't need to calculate additional correlation value because we already computed p value
#we must however also consider fold change. so we will make volcano plots for each of these comparisons
#to do this we first need to compute fold change

#each row (row not column ı checked)
#log2FC_gene1 = mean(condition1_gene) -mean(condition2_gene) @wanjun

log2FC1 = apply(BRCAtum, 1, mean) - apply(KIRCtum, 1, mean)

log2FC2 = apply(BRCAtum, 1, mean) - apply(LUADtum, 1, mean)

log2FC3 = apply(BRCAtum, 1, mean) - apply(PRADtum, 1, mean)

log2FC4 = apply(BRCAtum, 1, mean) - apply(THCAtum, 1, mean)

#PROBLEM: NaN's get produced here (PLS DEBUG)

```


```{r Vocano, with higher threshold }

#we make the volcano plots for all cancer types

df_volcanoplotBRCAvsKIRC = data.frame (log2FC1, p_BRCAvsKIRC)
df_volcanoplotBRCAvsKIRC$different = "NO"

df_volcanoplotBRCAvsKIRC$different[df_volcanoplotBRCAvsKIRC$log2FC1 > 0.1  &     df_volcanoplotBRCAvsKIRC$p_BRCAvsKIRC < 0.0000000000000000000001]= "UP"

df_volcanoplotBRCAvsKIRC$different[df_volcanoplotBRCAvsKIRC$log2FC1 < 0.1 &      df_volcanoplotBRCAvsKIRC$p_BRCAvsKIRC < 0.0000000000000000000001]= "DOWN"

#we plot 
library(ggplot2)
ggplot (data = df_volcanoplotBRCAvsKIRC, aes (x=log2FC1, y=-log10(p_BRCAvsKIRC), col = different)) + ggtitle("volcanoplot BRCA tumour vs KIRC tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )




df_volcanoplotBRCAvsLUAD = data.frame (log2FC2, p_BRCAvsLUAD)
df_volcanoplotBRCAvsLUAD$different = "NO"

df_volcanoplotBRCAvsLUAD$different[df_volcanoplotBRCAvsLUAD$log2FC2 > 0.1  &     df_volcanoplotBRCAvsLUAD$p_BRCAvsLUAD < 0.0000000000000000000001]= "UP"

df_volcanoplotBRCAvsLUAD$different[df_volcanoplotBRCAvsLUAD$log2FC2 < 0.1 &      df_volcanoplotBRCAvsLUAD$p_BRCAvsLUAD < 0.0000000000000000000001]= "DOWN"

#we plot

ggplot (data = df_volcanoplotBRCAvsLUAD, aes (x=log2FC2, y=-log10(p_BRCAvsLUAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs LUAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )




df_volcanoplotBRCAvsPRAD = data.frame (log2FC3, p_BRCAvsPRAD)
df_volcanoplotBRCAvsPRAD$different = "NO"

df_volcanoplotBRCAvsPRAD$different[df_volcanoplotBRCAvsPRAD$log2FC3 > 0.1  &     df_volcanoplotBRCAvsPRAD$p_BRCAvsPRAD < 0.0000000000000000000001]= "UP"

df_volcanoplotBRCAvsPRAD$different[df_volcanoplotBRCAvsPRAD$log2FC3 < 0.1 &      df_volcanoplotBRCAvsPRAD$p_BRCAvsPRAD < 0.0000000000000000000001]= "DOWN"

#we plot 
library(ggplot2)
ggplot (data = df_volcanoplotBRCAvsPRAD, aes (x=log2FC3, y=-log10(p_BRCAvsPRAD), col = different)) + ggtitle("volcanoplot BRCA tumour vs PRAD tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )




df_volcanoplotBRCAvsTHCA = data.frame (log2FC4, p_BRCAvsTHCA)
df_volcanoplotBRCAvsTHCA$different = "NO"

df_volcanoplotBRCAvsTHCA$different[df_volcanoplotBRCAvsTHCA$log2FC4 > 0.1  &     df_volcanoplotBRCAvsTHCA$p_BRCAvsTHCA < 0.0000000000000000000001]= "UP"

df_volcanoplotBRCAvsTHCA$different[df_volcanoplotBRCAvsTHCA$flog2FC4 < 0.1 &      df_volcanoplotBRCAvsTHCA$p_BRCAvsTHCA < 0.0000000000000000000001]= "DOWN"

#we plot 
library(ggplot2)
ggplot (data = df_volcanoplotBRCAvsTHCA, aes (x=log2FC4, y=-log10(p_BRCAvsTHCA), col = different)) + ggtitle("volcanoplot BRCA tumour vs THCA tumour") + geom_point() + theme_minimal() + scale_color_manual(values = c ("pink","black", "lightblue"))+ theme (text=element_text(size = 8) )


```










FURTHER DİAGNOSTİC TESTS and DEBUGGİNG (can delete if wanted. not necessary for function)
```{r compute p-values for each genes}
#DİAGNOSTİK TESTS

View(BRCAtum)
View(KIRCtum)

nrow(BRCAtum) == nrow(KIRCtum) # returns true
# patients are different but the genes are not

match(rownames(BRCAtum), rownames(KIRCtum))
#reihenfolge of the genes are also the same. we can get to work now

length(BRCAtum[1,]) == length(KIRCtum[1,]) # returns false v sad
length(BRCAtum[1,]) - length(KIRCtum[1,]) #BRCA has 39 more patients

#to run wilcox test we need the patient numbers to be same
min( length(BRCAtum[1,]), length(KIRCtum[1,]) , length(LUADtum[1,]) , length(PRADtum[1,]) , length(THCAtum[1,]) )
max( length(BRCAtum[1,]), length(KIRCtum[1,]) , length(LUADtum[1,]) , length(PRADtum[1,]) , length(THCAtum[1,]) )

#max 111 min 52. we need to exclude a lot of patients. who we gonna exclude? 
#no wait it do b working when paired = false


```


