---
title: "DataCleaningCompleate"
author: "Ekin Ören"
date: '2022-05-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


BIOTYPE FİLTERİNG BİG BOY:

_________________________________________________________________________________________________________________________________________________
GİRLİES REMEMBER. YOU NEED TO HAVE ALL THE DATABASES AND THE COPİES AS WELL AS ALL THE LİBRARİES FROM "DataLoadAndRead" FOR THİS CODE TO WORK FOR YOU. SO GO RUN THAT AND MAKE SURE THOSE MF's BE FLOATİNG İN YOUR ENVİROMENT! LOVE YOUU!

Errinnerung: we are working on the copies of the datasets, defined by the code below from last time.

CancerGeneExpressionsRemix = tcga_exp
CancerGeneExpressionsRemixAnnots = tcga_annot
TumorSpesificGenexpDatalist = tcga_tumor_norm
CancerHallmarkGenesetsList = genesets


This Chunk Configures Biomart, and gets the Gene names sepearately out of our data

```{r ConfigureBiomart, echo=FALSE}
#-----------------------------------------------------------#
# Load the genes
#-----------------------------------------------------------#

#directory = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/CancerGeneExpressionRemix.rds"
#filename = "CancerGeneExpressionRemix"
#genes <- readRDS(directory)
#help("readRDS")

tcga_exp = readRDS("C:/Users/EKN/Documents/tcga_tumor_log2TPM.RDS")
directory = "C:/Users/EKN/Documents/tcga_tumor_log2TPM.RDS"
genes <- readRDS(directory)
GeneNames <- rownames(genes)

#-----------------------------------------------------------#
# Split ENSG identifier from Gene Symbol
#-----------------------------------------------------------#


gene_identifier <- strsplit(GeneNames, split = "|", fixed = TRUE)


genes_ens<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[1]
  return(ens)
})

genes_symbol<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[2]
  return(ens)
})


# remove version number

genes_ens_vn <- strsplit(genes_ens, split = ".", fixed = TRUE)
genes_ensembl <- sapply(genes_ens_vn, FUN = function(genes_ens_vn){
  ens = genes_ens_vn[1]
  return(ens)
})

```
now we have a list of ENSEMBL-ID's for our genes under the name genes_ensembl



This chunk İdentifies the Biotypes of the genes we have

```{r chunky, echo=FALSE}
library(biomaRt)


# Access Database and use homo sapiens dataset

#ensembl = useEnsembl(biomart = "ensembl")   #this line does not work. SKİPPED İT
searchDatasets(mart = ensembl, pattern = "hsapiens")                              
ensembl <- useDataset(dataset = "hsapiens_gene_ensembl", mart = ensembl) 

# Search of gene biotypes using Ensemble Gene IDs as identifier
list_biotype = getBM(attributes = c("ensembl_gene_id", "gene_biotype"), filters = "ensembl_gene_id",
                     values = genes_ensembl, mart = ensembl)                                                                  

# Genes which could not be found in the database
rest = genes_ensembl[-match(list_biotype$ensembl_gene_id, genes_ensembl)]

```
list_biotype: The biotypes of the genes from our dataset (bigboi) genes that are found on ensembl database
rest: the genes from our database that doesnt exist



This chunk İdentifies the Biotypes found in known Hallmark Genes

```{r chunky, echo=FALSE}

## initiate package
if(!require(msigdbr))
  install.packages("msigdbr")
library(msigdbr)
library(biomaRt)

## retrieve hallmark datasets from msigdbr database. 

h_gene_sets = msigdbr(species = "Homo sapiens", category = "H")

#View(h_gene_sets)

## create an vector consist of ensemble (gene id) from the h_gene_sets 

hallmark_gen_id = c(h_gene_sets$ensembl_gene)

#View(hallmark_gen_id)

## Access Database and use homo sapiens dataset

ensembl = useEnsembl(biomart = "ensembl")   
searchDatasets(mart = ensembl, pattern = "hsapiens")                              
ensembl <- useDataset(dataset = "hsapiens_gene_ensembl", mart = ensembl) 

## Search of gene biotypes using Ensemble Gene IDs as identifier

list_biotype_hallmark = getBM(attributes = c("ensembl_gene_id", "gene_biotype"), filters = "ensembl_gene_id", values = hallmark_gen_id, mart = ensembl)                                       


View(list_biotype_hallmark)

## See what for biotypes are there in this datasets

cancery_biotypes = unique (list_biotype_hallmark$gene_biotype)
cancery_biotypes
#

#We decide NOT TO use the cancery biotypes but only use the protein coding ones!!!

```
hallmark_gen_id: ID's of hallmark genes
list_biotype_hallmark: list of hallmark gen ID's with biotypes
cancery_biotypes: biotypes found in hallmark genes


This Chunk finds the genes with the cancery biotypes.

```{r chunky, echo=FALSE}

protein_coding = list_biotype[which(list_biotype$gene_biotype == "protein_coding"), 1]

polymorphic_pseudogene = list_biotype[which(list_biotype$gene_biotype == "polymorphic_pseudogene"), 1]

lncRNA = list_biotype[which(list_biotype$gene_biotype == "lncRNA"), 1]


processed_pseudogene = list_biotype[which(list_biotype$gene_biotype == "processed_pseudogene"), 1]

transcribed_unprocessed_pseudogene = list_biotype[which(list_biotype$gene_biotype == "transcribed_unprocessed_pseudogene"),1]

interesting_ens = c(protein_coding, polymorphic_pseudogene, lncRNA, processed_pseudogene, transcribed_unprocessed_pseudogene)

head(interesting_ens)

```
interesting_ens: names of all our genes from relevant biotypes


This Chunk Pulls only the relevant data and makes a smaller data frame "SkinnyBoy"
```{r chunky, echo=FALSE}

directory = "C:/Users/EKN/Documents/tcga_tumor_log2TPM.RDS"
BigBoy <- readRDS(directory)
head(BigBoy)

matchingGetSkinny = match(protein_coding, genes_ensembl)
matchingGetSkinny

SkinnyBoy = BigBoy[matchingGetSkinny,]
dim(BigBoy)
dim(SkinnyBoy)
head(SkinnyBoy)

BiotypeFilteredBoy = SkinnyBoy

# Take only the matches.
#final_genes = interesting_ens[na.omit(matching)]
#final_genes


```
SkinnyBoy: The reduced data frame with only the relevant genes


__________________________________________________________________________________________________________________________________________________

NOW VARİANCE FİLTERİNG:


Select rows where variance is bigger than 75 percent (Interesting values)
```{r cars}

#Variance 
varianceSkinnyBoy = apply(BiotypeFilteredBoy, 1, var)
#greater75.variance = BiotypeFilteredBoy[varianceSkinnyBoy > quantile(varianceSkinnyBoy , probs = 0.95), ]

SkinnyBoy = BiotypeFilteredBoy[varianceSkinnyBoy > quantile(varianceSkinnyBoy , probs = 0.25), ]


```
greater75.varience: A data frame with unwanted rows


Find row numbers for greater75.varience rows.
```{r cars}

#matching.2 = match(rownames(greater75.variance), rownames(BiotypeFilteredBoy))
#matching.2 # a vector with the row numbers of Bigboy Dataset that needs to be removed

```
matching.2: row numbers in BigBoy we dont want because they are outliers



This chunk finds Unwanted Rows. (identifies rownumbers of the unwanted rows of bigboy using previous data)
```{r cars}

#dim(BiotypeFilteredBoy)
#Vector = c(1:19566) #A vector with all the row numbers of SkinnyBoy
#SkinnyBoy = BiotypeFilteredBoy[Vector[-matching.2],]

SkinnyBoy[1:10,]
dim(BiotypeFilteredBoy)
dim(SkinnyBoy)

saveRDS(SkinnyBoy, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/SkinnyBoy.RDS") 
#SkinnyBoy

```
__________________________________________________________________________________________________________________________________________________


SAME CODE FOR THE BREAST CANCER DATA:


_________________________________________________________________________________________________________________________________________________


CancerGeneExpressionsRemix = tcga_exp
CancerGeneExpressionsRemixAnnots = tcga_annot
TumorSpesificGenexpDatalist = tcga_tumor_norm
CancerHallmarkGenesetsList = genesets


This Chunk Configures Biomart, and gets the Gene names sepearately out of our data

```{r ConfigureBiomart, echo=FALSE}
#-----------------------------------------------------------#
# Load the genes
#-----------------------------------------------------------#

#directory = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/CancerGeneExpressionRemix.rds"
#filename = "CancerGeneExpressionRemix"
#genes <- readRDS(directory)
#help("readRDS")

tcga_tumor_norm = readRDS("C:/Users/EKN/Documents/tcga_tumor_normal_datascience_proj_2022.RDS")
BigSadTT = tcga_tumor_norm$BRCA$tumor
BigHappyTT = tcga_tumor_norm$BRCA$normal

head(tcga_tumor_norm$BRCA$tumor)
head(BigSadTT)

genes1 <- BigSadTT
genes2 <- BigHappyTT
GeneNames1 <- rownames(genes1)
GeneNames2 <- rownames(genes2)
GeneNames1[1:10]

#-----------------------------------------------------------#
# Split ENSG identifier from Gene Symbol
#-----------------------------------------------------------#


gene_identifier1 <- strsplit(GeneNames1, split = "|", fixed = TRUE)


genes_ens1<- sapply(gene_identifier1, FUN = function(gene_identifier1){
  ens = gene_identifier1[1]
  return(ens)
})

genes_symbol1<- sapply(gene_identifier1, FUN = function(gene_identifier1){
  ens = gene_identifier1[2]
  return(ens)
})


gene_identifier2 <- strsplit(GeneNames2, split = "|", fixed = TRUE)


genes_ens2<- sapply(gene_identifier2, FUN = function(gene_identifier2){
  ens = gene_identifier2[1]
  return(ens)
})

genes_symbol2<- sapply(gene_identifier2, FUN = function(gene_identifier2){
  ens = gene_identifier2[2]
  return(ens)
})



# remove version number

genes_ens_vn1 <- strsplit(genes_ens1, split = ".", fixed = TRUE)
genes_ensembl1 <- sapply(genes_ens_vn1, FUN = function(genes_ens_vn1){
  ens = genes_ens_vn1[1]
  return(ens)
})


genes_ens_vn2 <- strsplit(genes_ens2, split = ".", fixed = TRUE)
genes_ensembl2 <- sapply(genes_ens_vn2, FUN = function(genes_ens_vn2){
  ens = genes_ens_vn2[1]
  return(ens)
})

genes_ensembl2[1:10]
genes_ens_vn2[1:10]

```
now we have a list of ENSEMBL-ID's for our genes under the names genes_ensembl1 (BigSadTT) and genes_ensembl2 (BigHappyTT)



This chunk İdentifies the Biotypes of the genes we have

```{r chunky, echo=FALSE}

## initiate package
if(!require(msigdbr))
  install.packages("msigdbr")
library(msigdbr)
library(biomaRt)

## retrieve hallmark datasets from msigdbr database. 

h_gene_sets = msigdbr(species = "Homo sapiens", category = "H")

## create an vector consist of ensemble (gene id) from the h_gene_sets 

hallmark_gen_id = c(h_gene_sets$ensembl_gene)

## Access Database and use homo sapiens dataset

ensembl = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/ensembl.RDS")
#ensembl = useEnsembl(biomart = "ensembl")
searchDatasets(mart = ensembl, pattern = "hsapiens")                              
ensembl <- useDataset(dataset = "hsapiens_gene_ensembl", mart = ensembl) 


# Search of gene biotypes using Ensemble Gene IDs as identifier
list_biotype1 = getBM(attributes = c("ensembl_gene_id", "gene_biotype"), filters = "ensembl_gene_id",
                     values = genes_ensembl1, mart = ensembl)                                                                  

# Genes which could not be found in the database
#rest1 = genes_ensembl[-match(list_biotype$ensembl_gene_id, genes_ensembl1)] #GİVES ERROR IDK WHY

# Search of gene biotypes using Ensemble Gene IDs as identifier
list_biotype2 = getBM(attributes = c("ensembl_gene_id", "gene_biotype"), filters = "ensembl_gene_id",
                     values = genes_ensembl2, mart = ensembl)                                                                  

# Genes which could not be found in the database
#rest2 = genes_ensembl[-match(list_biotype$ensembl_gene_id, genes_ensembl2)] #GİVES ERROR IDK WHY

```
list_biotype1 and 2: The biotypes of the genes from our datasets (BigSadTT and HappyTT) genes that are found on ensembl database
rest1 and 2: the genes from our datasets that doesnt exist in the database



For comparison we use the cancery biotypes that are previously detected

```{r chunky, echo=FALSE}

cancery_biotypes = unique (list_biotype_hallmark$gene_biotype)
cancery_biotypes

```
hallmark_gen_id: ID's of hallmark genes
list_biotype_hallmark: list of hallmark gen ID's with biotypes
cancery_biotypes: biotypes found in hallmark genes


This Chunk finds the genes with the cancery biotypes.

```{r chunky, echo=FALSE}

protein_coding1 = list_biotype1[which(list_biotype1$gene_biotype == "protein_coding"), 1]

polymorphic_pseudogene1 = list_biotype1[which(list_biotype1$gene_biotype == "polymorphic_pseudogene"), 1]

lncRNA1 = list_biotype1[which(list_biotype1$gene_biotype == "lncRNA"), 1]


processed_pseudogene1 = list_biotype1[which(list_biotype1$gene_biotype == "processed_pseudogene"), 1]

transcribed_unprocessed_pseudogene1 = list_biotype1[which(list_biotype1$gene_biotype == "transcribed_unprocessed_pseudogene"),1]

interesting_ens1 = c(protein_coding1, polymorphic_pseudogene1, lncRNA1, processed_pseudogene1, transcribed_unprocessed_pseudogene1)

head(interesting_ens1)



protein_coding2 = list_biotype2[which(list_biotype2$gene_biotype == "protein_coding"), 1]

polymorphic_pseudogene2 = list_biotype2[which(list_biotype2$gene_biotype == "polymorphic_pseudogene"), 1]

lncRNA2 = list_biotype2[which(list_biotype2$gene_biotype == "lncRNA"), 1]


processed_pseudogene2 = list_biotype2[which(list_biotype2$gene_biotype == "processed_pseudogene"), 1]

transcribed_unprocessed_pseudogene2 = list_biotype2[which(list_biotype2$gene_biotype == "transcribed_unprocessed_pseudogene"),1]

interesting_ens2 = c(protein_coding2, polymorphic_pseudogene2, lncRNA2, processed_pseudogene2, transcribed_unprocessed_pseudogene2)

head(interesting_ens2)
head(interesting_ens1)


```
interesting_ens: names of all our genes from relevant biotypes


This Chunk Pulls only the relevant data and makes a smaller data frame
```{r chunky, echo=FALSE}


matching1 = match(protein_coding1, genes_ensembl1)
matching2 = match(protein_coding2, genes_ensembl2)

length(BigSadTT[1,])
length(matching1)
length(unique(matching1))

BiotypeCleanedSadTT  = BigSadTT[matching1,]
BiotypeCleanedHappyTT = BigHappyTT[matching2,]

BigSadTT[1,]


dim(BigSadTT)
dim(BigHappyTT)
dim(SkinnySadTT)
dim(SkinnyHappyTT)
# Take only the matches.
#final_genes = interesting_ens[na.omit(matching)]
#final_genes


```
SkinnyBoy: The reduced data frame with only the relevant genes


_____________________________________________________________________________________________________________________________________


NOW VARİANCE FİLTERİNG:


Finished varience filtering for TT datasets in this chunk
```{r cars}

#Variance 
variance1 = apply(BiotypeCleanedSadTT, 1, var)
variance2 = apply(BiotypeCleanedHappyTT, 1, var)

greater75.variance1 = SkinnySadTT[variance1 > quantile(variance1 , probs = 0.05), ]
greater75.variance2 = SkinnyHappyTT[variance2 > quantile(variance2, probs = 0.05), ]

dim(greater75.variance1)

SkinnySadTT = greater75.variance1
SkinnyHappyTT = greater75.variance2

saveRDS(SkinnySadTT, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/SkinnySadTT.RDS") 
saveRDS(SkinnyHappyTT, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/SkinnyHappyTT.RDS") 

```


READ THE COMMENTS BELOW! DO NOT RUN WİTHOUT READİNG. 
READ THE COMMENTS BELOW! DO NOT RUN WİTHOUT READİNG.
READ THE COMMENTS BELOW! DO NOT RUN WİTHOUT READİNG.
READ THE COMMENTS BELOW! DO NOT RUN WİTHOUT READİNG.
READ THE COMMENTS BELOW! DO NOT RUN WİTHOUT READİNG.

_____________________________________________________________________________________________________________________________________
ONLY JUNK ROWS BELOW DO NOT FUCKİNG RUN THEM
_____________________________________________________________________________________________________________________________________
Select rows where variance is bigger than 75 percent (Interesting values)
```{r cars}

#Variance 
varianceSkinnyBoy = apply(BiotypeFilteredBoy, 1, var)
#greater75.variance = BiotypeFilteredBoy[varianceSkinnyBoy > quantile(varianceSkinnyBoy , probs = 0.95), ]

SkinnyBoy = BiotypeFilteredBoy[varianceSkinnyBoy > quantile(varianceSkinnyBoy , probs = 0.25), ]


```




Find row numbers for greater75.varience rows.
```{r cars}

matchingBoobs1 = match(rownames(greater75.variance1), rownames(SkinnySadTT))
matchingBoobs1 # a vector with the row numbers of Bigboy Dataset that needs to be removed


matchingBoobs2 = match(rownames(greater75.variance2), rownames(SkinnyHappyTT))
matchingBoobs2 # a vector with the row numbers of Bigboy Dataset that needs to be removed

```
matching.2: row numbers in BigBoy we dont want because they are outliers



This chunk finds Unwanted Rows. (identifies rownumbers of the unwanted rows of bigboy using previous data)
```{r cars}
dim(SkinnySadTT)
dim(SkinnyHappyTT)

Vector = c(1:19306) #A vector with all the row numbers of SkinnyBoy
SkinnySadTT = SkinnySadTT[Vector[-matchingBoobs1],]
SkinnyHappyTT = SkinnyHappyTT[Vector[-matchingBoobs2],]


dim(SkinnySadTT)
dim(SkinnyHappyTT)

```
___________________________________________________________________________________________________________________________________


This chunk deletes all new stuff so you can run it again if you have doubts
```{r cars}

rm(SkinnyBoy)
rm(SkinnyHappyTT)
rm(SkinnySadTT)

```


This chunk is for tries
```{r cars}

vectorA = c("gene one", "gene two", "gene three")
VectorB = c("gene one", "gene three")

vectorA

match(vectorA,VectorB)
match(VectorB,vectorA)

```






















