---
title: "Fixing_Pathway_Matrix"
author: "Ekin Ören"
date: '2022-07-01'
output: html_document
---


Here are the GSVA relevant libraries to load if you dont have them already
```{r hehehe ,}

#------------------------------------------------------------------------------#
# Geneset Variation Analysis (GSVA) Demo [Hänzelmann et al., 2013]
#-----------------------------------------------------------------------------#

if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("GSVA")

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("BiocParallel")

library(GSVA)
library(BiocParallel)
library(msigdbr)
library(ggplot2)
library(uwot)

```



Get CPathways and match them to ensembl IDS
```{r hehehe ,}


library("biomaRt")


ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")



CPathways = list ("ID")

for (i in 6:6366){
   CPathways$ID[i] =   getBM(attributes=c("ensembl_gene_id") ,mart=ensembl,
  filters= "hgnc_symbol" , values = h_pathways[i])
} 

 
View(CPathways)

CPathways = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/CPathways.RDS")

CPathwaysNames = names(h_pathways)

saveRDS(CPathwaysNames, "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/CPathwaysNames.RDS")


## attributes: what we wanna retrieve from Biomart 
## value: our datasets -> retrieve ensembl_gene_id from our datasets 

```


Get expressions and embed p values into the pathway matrix for all pathways EXAMPLE CODE AND EXPLANATİONS
```{r pea time ,}

SkinnyBoyEnsembl
View(x)

pathway_geneexp_patient_SB = list ()
pathway_geneexp_patient_SB [[1]]= SkinnyBoy[match(unlist(x$ID[1]), SkinnyBoyEnsembl), 1] #gets the matching genes
 banana6= -match(unlist(x$ID[1]),SkinnyBoyEnsembl)
  banana7 <- banana6[!is.na(banana6)] #banana 7 is the ensembl ıds of our pathway with the NA's removed and minus in front
  non_pathway_geneexp_patient1_vector_SB = SkinnyBoy[banana7,1] #gets all genes BUT the pathway
t.test(pathway_geneexp_patient_SB[[1]], non_pathway_geneexp_patient1_vector_SB)$p.value 

```



Get expressions and embed p values into the pathway matrix for Metabolic pathways
```{r pea time ,}

for (m in 1:9741){
  for (n in 1:46){
      pathway_geneexp_patient_SB[[n]]= SkinnyBoy[match(unlist(x$ID[n]),genes_ensembl_SB),m]
     banana = -match(unlist(x$ID[n]),genes_ensembl_SB)
     banana <- banana[!is.na(banana)]
     non_pathway_geneexp_patient_vector_SB = SkinnyBoy[banana,m]
     df_pathway_patient_SB[n,m] = t.test(pathway_geneexp_patient_SB[[n]], non_pathway_geneexp_patient_vector_SB)$p.value
  }
}



view(x) #list of hallmark ensembls
View(HallmarksGenesets) #list of hallmark pathways

View(MetabolicPatEnsembl) #list of metabolic ensembles
View(MetabolicGenesets) #lisr of metabolic pathways

View(CPathways) #list of CPathway ensembles 
View(h_pathways) #list of CPathway pathways


MetabolicGenesets = readRDS(file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/GeneNamesMetabolicPatsLit.rds")




MetabolicPathway_geneexp_list = list()
df_MetPathway = data.frame()

for (m in 1:9741){
  for (n in 1:46){
      MetabolicPathway_geneexp_list[[n]]= SkinnyBoy[match(unlist(MetabolicsEnsembl$ID[n]),genes_ensembl_SB),m]
     banana = -match(unlist(MetabolicsEnsembl$ID[n]),genes_ensembl_SB)
     banana <- banana[!is.na(banana)]
     non_pathway_geneexp_patient_vector_SB = SkinnyBoy[banana,m]
     df_MetPathway[n,m] = t.test(MetabolicPathway_geneexp_list[[n]], non_pathway_geneexp_patient_vector_SB)$p.value
    
  }
}



#this lil group puts pathway names where they belong
back_pathN = c()

  for (n in 1:46){
    
    back_pathNhuoo = append(back_pathN,names(MetabolicGenesets[[n]]))

  }

back_pathN[which(duplicated(back_pathN))]= paste(back_pathN[which(duplicated(back_pathN))], "duplicate")
rownames(df_MetPathway) = back_pathN


#now we add patient names

humans = colnames(SkinnyBoy)
colnames(df_MetPathway) = colnames(SkinnyBoy)


View(df_MetPathway)

```



Get expressions and do GSVA scoring thus making the final pathway matrix for Metabolic pathways
```{r pea time ,}

#first lets merge these motherfuckers

View(x) #list of hallmark ensembls
View(HallmarksGenesets) #list of hallmark pathways

View(MetabolicPatEnsembl) #list of metabolic ensembles
View(MetabolicGenesets) #lisr of metabolic pathways

View(C_PathwaysEns) #list of CPathway ensembles DOESNT WORK MUST REDO
View(h_pathways) #list of CPathway pathways
#####################################################################################################

Metabolic.names = c()

for (i in 1:46){
  Metabolic.names [i] = names (MetabolicGenesets[[i]]) 
}

saveRDS(Metabolic.names,"C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/MetabolicPathwayNames.rds" )
        
        
        
        ##################################MAKES MIX OF ALL PATHWAYS İN ORDER Hall Met C############

ha = x$ID
huaa = CPathways$ID
View(huaa)

AllPathwaysEns_List = c(ha, MetabolicPatEnsembl$ID, huaa )

View(AllPathwaysEns_List)


AllPathwayNames = c()
View(AllPathwayNames)

AllPathwayNames = append(AllPathwayNames,names(HallmarksGenesets[[1]]))


  for (n in 1:46){
    
    AllPathwayNames = append(AllPathwayNames,names(MetabolicGenesets[[n]]))

  }

  for (n in 1:6366){
    
    AllPathwayNames = append(AllPathwayNames,names(h_pathways[n]))

  }

AllPathwayNames[which(duplicated(AllPathwayNames))]= paste(AllPathwayNames[which(duplicated(AllPathwayNames))], "_isduplicate")
################################################################################################



#THIS CODE PİECE NAMES THE PATHWAYS###########################################

AllPathwaysEns_List = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/AllPathwaysEns_List.rds")

View(AllPathwaysEns_List)

names(AllPathwaysEns_List) <- AllPathwayNames

saveRDS(AllPathwaysEns_List, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/AllPathwaysEns_List.rds")

################################################################################################

#Updates the names of skinnyboy
row.names(SkinnyBoy) = SkinnyBoyEnsembl
saveRDS(SkinnyBoy, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/SkinnyBoy.RDS")

```



This chunk runs GSVA and returns our Final Pathway Matrix

```{r chonk time ,}

## die gsva funktion erwartet den ganzen dataframe mit allen tumortypes, um das gene ranking zu machen (siehe abbildung Figure1)
## dasselbe gilt für normal vs tumor (macht einen df und schmeißt den rein)

AllPathwaysEns_List #a list of pathway ensembl IDS
SkinnyBoy #a data frame with ENSEMBL ID as rowname and expression data

library(GSVA)
library(BiocParallel)
library(msigdbr)
library(ggplot2)
library(uwot)

params <- MulticoreParam(workers = multicoreWorkers())

pw_activity <- gsva(expr = as.matrix(SkinnyBoy),
                    gset.idx.list = AllPathwaysEns_List,
                    kcdf = "Gaussian",
                    min.sz = 15,
                    BPPARAM = params)



saveRDS(pw_activity, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/PathwayMatrixDeluxePremium.rds")


```



for running diagnostics (you can delete if wanted)

```{r chonk time ,}

#first make sure SB ensembl IDS are fixed

row.names(SkinnyBoy) = SkinnyBoyEnsembl
View(SkinnyBoy)

saveRDS(SkinnyBoy, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/SkinnyBoy.RDS")
SkinnyBoy = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/SkinnyBoy.RDS")
View(SkinnyBoy)


#final diagnostics
match(names(AllPathwaysEns_List),rownames(pw_activity))

pw_activity[1000,1:5]

rownames(pw_activity[600:700,])

```





