---
title: "Metabolic_Pathway_Genes_Puller"
author: "Ekin Ören"
date: '2022-06-05'
output: html_document
---


LOAD junona's List of gene names per pathway
```{r cathatcat, echo=FALSE}

GeneNamesMetabolicPatsLit = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/lit.pathways.l.RDS.rds")
View(GeneNamesMetabolicPatsLit)
GeneNamesMetabolicPatsLit[1]
```

Pull gene IDs of each gene from ensembl.
```{r cathatcat, echo=FALSE}

library("biomaRt")

ensembl = useMart("ensembl",dataset="hsapiens_gene_ensembl")

MetabolicsEnsembl = list ("ID")

for (i in 1:46){
  
   MetabolicsEnsembl$ID[i] =   getBM(attributes=c("ensembl_gene_id") ,mart=ensembl,
  filters= "hgnc_symbol" , values = GeneNamesMetabolicPatsLit[[i]])
} 

View(MetabolicsEnsembl)

## attributes: what we wanna retrieve from Biomart 
## value: our datasets -> retrieve ensembl_gene_id from our datasets 
## i have here the pathways (all) and all the ensemble ID related to the respectively pathway

```

```{r For june to load genes_ensembl_SB}
# since biomart no worky 
readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/genes_ensembl_SB.RDS")
readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/GeneNamesMetabolicPatsLit.RDS")
readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/genes_ensembl_SB.RDS")
readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/MetabolicsEnsembl.RDS")#this wrong this empty
readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/MetabolicEnsembl.rds")
```

Get geneID's from skinnyboy and save them as a vector
```{r cathatcat, echo= TRUE}


GeneNamesSB <- rownames(SkinnyBoy)


gene_identifierSB <- strsplit(GeneNamesSB, split = "|", fixed = TRUE)

genes_ens_SB<- sapply(gene_identifierSB, FUN = function(gene_identifierSB){
  ens = gene_identifierSB[1]
  return(ens)
})

genes_symbol_SB<- sapply(gene_identifierSB, FUN = function(gene_identifierSB){
  ens = gene_identifierSB[2]
  return(ens)
})

# remove version number

genes_ens_vn_SB <- strsplit(genes_ens_SB, split = ".", fixed = TRUE)
genes_ensembl_SB <- sapply(genes_ens_vn_SB, FUN = function(genes_ens_vn_SB){
  ens = genes_ens_vn_SB[1]
  return(ens)
})
######


genes_ens_SB[1:10]
#saveRDS (genes_ensembl_SB, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/genes_ens_SB.RDS")

```


TESTİNG THE CODE FOR 10 PATİENTS FİRST
```{r cathatcat, echo=TRUE}
MetabolicPathway_geneexp_list = list ()
df_MetPathwayFixed = data.frame()
df_MetPathway = data.frame ()
View(SkinnyBoy)

for (m in 1:ncol(SkinnyBoy)){
  for (n in 1:46){
      MetabolicPathway_geneexp_list[[n]]= SkinnyBoy[match(unlist(MetabolicsEnsembl$ID[n]),genes_ensembl_SB),m]
     banana = -match(unlist(MetabolicsEnsembl$ID[n]),genes_ensembl_SB)
     banana <- banana[!is.na(banana)]
     non_pathway_geneexp_patient_vector_SB = SkinnyBoy[banana,m]
     df_MetPathway[n,m] = wilcox.test(MetabolicPathway_geneexp_list[[n]], non_pathway_geneexp_patient_vector_SB)$p.value
 df_MetPathwayFixed[n,m] = df_MetPathway[n,m] * number_pathway[n]    
  }
}

View (df_MetPathwayFixed)
number_pathway = c()
df_MetPathwayFixed = data.frame()

number_pathway = list ()

for (i in 1:46) {
  number_pathway[i]= length(unlist(MetabolicEnsembl$ID[[i]]))
}

for (m in 1:ncol (SkinnyBoy)){
  for (n in 1:46) {
  df_MetPathwayFixed[n,m] = unlist(number_pathway[n]) * unlist(df_MetPathway[n,m])
  }
}


View(df_MetPathway)
View(df_MetPathwayFixed)
View(number_pathway)


```


Create the Final df_MetPathway which is a DF of p values per patient for 46 metabolic pathways
```{r cathatcat, echo=TRUE}


MetabolicPathway_geneexp_list = list()
df_MetPathway = data.frame()


for (m in 1:9741){
  for (n in 1:46){
      MetabolicPathway_geneexp_list[[n]]= SkinnyBoy[match(unlist(MetabolicsEnsembl$ID[n]),genes_ensembl_SB),m]
     banana = -match(unlist(MetabolicsEnsembl$ID[n]),genes_ensembl_SB)
     banana <- banana[!is.na(banana)]
     non_pathway_geneexp_patient_vector_SB = SkinnyBoy[banana,m]
     df_MetPathway[n,m] = t.test(MetabolicPathway_geneexp_list[[n]], non_pathway_geneexp_patient_vector_SB)$p.value
     
  }
}



(df_MetPathway)

number_pathway = c()
df_MetPathwayFixed = data.frame()

for (m in 1:9741){
  for (n in 1:46) {
  number_pathway[n]= length(unlist(MetabolicsEnsembl$ID[[n]]))
  df_MetPathwayFixed[n,m] = df_MetPathway[n,m] * number_pathway[n]
    
  }
}

```


Views all pathway dataframes, adds them into a big main dataframe views that dataframe
```{r cathatcat, echo=TRUE}

MetabolicPathwayDF = data.frame()
Veccy1 <-rep( "Metabolic" ,times=length(df_MetPathwayFixed[,1]))
MetabolicPathwayDF = df_MetPathwayFixed
MetabolicPathwayDF$Type <- Veccy1
View(MetabolicPathwayDF)


HallmarkPathwayDF = data.frame()
Veccy2 <-rep( "Hallmark" ,times=length(df_pathway_patient_S_pvalue[,1]))
HallmarkPathwayDF = df_pathway_patient_S_pvalue
HallmarkPathwayDF$Type <- Veccy2
View(HallmarkPathwayDF)

CompleatePathwayDF = rbind(HallmarkPathwayDF, MetabolicPathwayDF)
View(CompleatePathwayDF)


```


```{r saving the final three DF}
saveRDS(MetabolicPathwayDF,file = "/Users/lunona/Desktop/BRCA/MetabolicPathwayDF.RDS")
saveRDS(HallmarkPathwayDF,file = "/Users/lunona/Desktop/BRCA/HallmarkPathwayDF.RDS")
saveRDS(CompleatePathwayDF,file ="/Users/lunona/Desktop/BRCA/CompleatePathwayDF.RDS")
```

THE DATA FRAMES WE GET AT THE END OF THİS MARKDOWN ARE LİSTED BELOW

MetabolicPathwayDF
HallmarkPathwayDF
CompleatePathwayDF





Debugging and diagnostics
```{r saving the final three DF}


AllPathways = data.frame()
A = c(1:92)
AllPathways[1,] = A

for (i in 1:9741)
{
  
  AllPathways[,i] = unlist(CompleatePathwayDF[[i]])
  
}

#AllPathways is Compleate Pathway DF without last col with MET and Hallmark


```


