---
title: "Venn_Diagramm"
author: "Ekin Ören"
date: '2022-07-01'
output: html_document
---


This chunk loads the library required for the venn diagram
```{r Venn Life Gets you down, echo=FALSE}

if (!require(devtools)) install.packages("devtools")
devtools::install_github("gaospecial/ggVennDiagram")

```



This chunk extracts genes and creates vectors that will be imputs into the venn diagramm

```{r Venn Life Gets you down, echo=FALSE}

#genes_ens_SB needs to be already in your enviroment! if not, ask me (echo) to send it

genes_ens_vn <- strsplit(genes_ens_SB, split = ".", fixed = TRUE)
SkinnyBoyEnsembl <- sapply(genes_ens_vn, FUN = function(genes_ens_vn){
  ens = genes_ens_vn[1]
  return(ens)
})


A = c()

for(i in 1:46)
{
  A= append(A, MetabolicsEnsembl$ID[[i]])
}

Metabolic_EnsemblesVector = unique(A)

Hallmarks = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/data/x.RDS")


B = c()

for(i in 1:46)
{
  B = append(B, Hallmarks$ID[[i]])
}

Hallmarks_EnsemblesVector = unique(B)

VoVee = list(Hallmarks_EnsemblesVector, Metabolic_EnsemblesVector, SkinnyBoyEnsembl)

```




This chunk computes the first Ven diagram of all pathway we had before addition of C compared
```{r Venn Life Gets you down, echo=FALSE}


AllPathwayGenesByID = append(Hallmarks_EnsemblesVector,Metabolic_EnsemblesVector)
VoVoo = list(AllPathwayGenesByID, SkinnyBoyEnsembl)

ggVennDiagram(
  VoVoo, label_alpha = 0,
  category.names = c("pathways","SkinnyBoy")
  ) +
  ggplot2::scale_fill_gradient(low="white",high = "violet")

```




This chunk computes the first Ven diagram of all pathway groups compared
```{r Venn Life Gets you down, echo=FALSE}


C = c()

for(i in 1:6366)
{
  C = append(C, CPathways$ID[[i]])
}

View(C)

CPathways = unique(C)


VoVouou = list(Hallmarks_EnsemblesVector, Metabolic_EnsemblesVector, SkinnyBoyEnsembl, CPathways)


ggVennDiagram(
  VoVouou, label_alpha = 0,
  category.names = c("Hallmarks","Metabolics","SkinnyBoy","CPathways")
  ) +
  ggplot2::scale_fill_gradient(low="white",high = "violet")


```




This chunk computes the Venn diagram of  pathway matrix selections and SkinnyBoy
```{r Venn Life Gets you down, echo=FALSE}

AllPathwayGenesByID2 = append(Hallmarks_EnsemblesVector,Metabolic_EnsemblesVector)
AllPathwayGenesByID3 = append(AllPathwayGenesByID2,CPathways)
VoVeh = list(AllPathwayGenesByID3, SkinnyBoyEnsembl)

ggVennDiagram(
  VoVeh, label_alpha = 0,
  category.names = c("pathways","SkinnyBoy")
  ) +
  ggplot2::scale_fill_gradient(low="white",high = "violet")

```


This chunk generates venn diagrams comparing old pathway matrix, new pathway matrix and the initial dataframe (SkinnyBoy)

```{r setup,}

A = c()

for (i in 1:6458)
{
  B = AllPathwaysEns_List[[i]]
  A = append(A, B)
}

EnsemblesVectorAllPathways = A

EnsemblesVectorSkinnyBoy = rownames(SkinnyBoy)

View(PathwayMatrixDeluxePremium)



for(i in 1:length(rownames(PathwayMatrixDeluxePremium)))
{
  apple1 = match(rownames(PathwayMatrixDeluxePremium), names(AllPathwaysEns_List))
  
}


for (i in apple1)
{
  apple4 = AllPathwaysEns_List[[i]]
}

EnsemblesVectorLastPathways = apple4





hes = list(EnsemblesVectorAllPathways, EnsemblesVectorSkinnyBoy, EnsemblesVectorLastPathways)

hus = list(unique(EnsemblesVectorAllPathways),unique(EnsemblesVectorSkinnyBoy),unique(EnsemblesVectorLastPathways))

library("ggVennDiagram")

ggVennDiagram(
  hus, label_alpha = 0,
  category.names = c("AllPathways","SkinnyBoy","Last Pathways")
  ) +
  ggplot2::scale_fill_gradient(low="white",high = "violet")





hunana = list(unique(EnsemblesVectorSkinnyBoy),unique(EnsemblesVectorLastPathways))

library("ggVennDiagram")

ggVennDiagram(
  hunana, label_alpha = 0,
  category.names = c("SkinnyBoy","Last Pathways")
  ) +
  ggplot2::scale_fill_gradient(low="white",high = "violet")


```


