---
title: "BiotypeFiltering"
author: "Ekin Ören"
date: '2022-05-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


GİRLİES REMEMBER. YOU NEED TO HAVE ALL THE DATABASES AND THE COPİES AS WELL AS ALL THE LİBRARİES FROM "DataLoadAndRead" FOR THİS CODE TO WORK FOR YOU. SO GO RUN THAT AND MAKE SURE THOSE MF's BE FLOATİNG İN YOUR ENVİROMENT! LOVE YOUU!

Errinnerung: we are working on the copies of the datasets, defined by the code below from last time.

CancerGeneExpressionsRemix = tcga_exp
CancerGeneExpressionsRemixAnnots = tcga_annot
TumorSpesificGenexpDatalist = tcga_tumor_norm
CancerHallmarkGenesetsList = genesets


This Chunk Configures Biomart, and gets the Gene names sepearately out of our data

```{r ConfigureBiomart, echo=FALSE}
#-----------------------------------------------------------#
# Load the genes
#-----------------------------------------------------------#

#directory = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/CancerGeneExpressionRemix.rds"
#filename = "CancerGeneExpressionRemix"
#genes <- readRDS(directory)
#help("readRDS")

tcga_exp = readRDS("C:/Users/EKN/Documents/tcga_tumor_log2TPM.RDS")
directory = "C:/Users/EKN/Documents/tcga_tumor_log2TPM.RDS"
genes <- readRDS(directory)
GeneNames <- rownames(genes)

#-----------------------------------------------------------#
# Split ENSG identifier from Gene Symbol
#-----------------------------------------------------------#


gene_identifier <- strsplit(GeneNames, split = "|", fixed = TRUE)


genes_ens<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[1]
  return(ens)
})

genes_symbol<- sapply(gene_identifier, FUN = function(gene_identifier){
  ens = gene_identifier[2]
  return(ens)
})


# remove version number

genes_ens_vn <- strsplit(genes_ens, split = ".", fixed = TRUE)
genes_ensembl <- sapply(genes_ens_vn, FUN = function(genes_ens_vn){
  ens = genes_ens_vn[1]
  return(ens)
})

View(genes_ensembl)

```
now we have a list of ENSEMBL-ID's for our genes under the name genes_ensembl



This chunk İdentifies the Biotypes of the genes we have

```{r chunky, echo=FALSE}
library(biomaRt)


# Access Database and use homo sapiens dataset
ensembl = useEnsembl(biomart = "ensembl")   
searchDatasets(mart = ensembl, pattern = "hsapiens")                              
ensembl <- useDataset(dataset = "hsapiens_gene_ensembl", mart = ensembl) 

# Search of gene biotypes using Ensemble Gene IDs as identifier
list_biotype = getBM(attributes = c("ensembl_gene_id", "gene_biotype"), filters = "ensembl_gene_id",
                     values = genes_ensembl, mart = ensembl)                                                                  

# Genes which could not be found in the database
rest = genes_ensembl[-match(list_biotype$ensembl_gene_id, genes_ensembl)]

```
list_biotype: The biotypes of the genes from our dataset (bigboi) genes that are found on ensembl database
rest: the genes from our database that doesnt exist



This chunk İdentifies the Biotypes found in known Hallmark Genes

```{r chunky, echo=FALSE}

## initiate package
if(!require(msigdbr))
  install.packages("msigdbr")
library(msigdbr)
library(biomaRt)

## retrieve hallmark datasets from msigdbr database. 

h_gene_sets = msigdbr(species = "Homo sapiens", category = "H")

#View(h_gene_sets)

## create an vector consist of ensemble (gene id) from the h_gene_sets 

hallmark_gen_id = c(h_gene_sets$ensembl_gene)

#View(hallmark_gen_id)

## Access Database and use homo sapiens dataset

ensembl = useEnsembl(biomart = "ensembl")   
searchDatasets(mart = ensembl, pattern = "hsapiens")                              
ensembl <- useDataset(dataset = "hsapiens_gene_ensembl", mart = ensembl) 

## Search of gene biotypes using Ensemble Gene IDs as identifier

list_biotype_hallmark = getBM(attributes = c("ensembl_gene_id", "gene_biotype"), filters = "ensembl_gene_id", values = hallmark_gen_id, mart = ensembl)                                       


View(list_biotype_hallmark)

## See what for biotypes are there in this datasets

cancery_biotypes = unique (list_biotype_hallmark$gene_biotype)
cancery_biotypes
#

```
hallmark_gen_id: ID's of hallmark genes
list_biotype_hallmark: list of hallmark gen ID's with biotypes
cancery_biotypes: biotypes found in hallmark genes


This Chunk finds the genes with the cancery biotypes.

```{r chunky, echo=FALSE}

protein_coding = list_biotype[which(list_biotype$gene_biotype == "protein_coding"), 1]

polymorphic_pseudogene = list_biotype[which(list_biotype$gene_biotype == "polymorphic_pseudogene"), 1]

lncRNA = list_biotype[which(list_biotype$gene_biotype == "lncRNA"), 1]


processed_pseudogene = list_biotype[which(list_biotype$gene_biotype == "processed_pseudogene"), 1]

transcribed_unprocessed_pseudogene = list_biotype[which(list_biotype$gene_biotype == "transcribed_unprocessed_pseudogene"),1]

interesting_ens = c(protein_coding, polymorphic_pseudogene, lncRNA, processed_pseudogene, transcribed_unprocessed_pseudogene)

head(interesting_ens)

```
interesting_ens: names of all our genes from relevant biotypes


This Chunk Pulls only the relevant data and makes a smaller data frame "SkinnyBoy"
```{r chunky, echo=FALSE}

directory = "C:/Users/EKN/Documents/tcga_tumor_log2TPM.RDS"
BigBoy <- readRDS(directory)
head(BigBoy)

matching = match(interesting_ens, genes_ensembl)
matching

SkinnyBoy = BigBoy[matching,]
dim(BigBoy)
dim(SkinnyBoy)
head(SkinnyBoy)
# Take only the matches.
#final_genes = interesting_ens[na.omit(matching)]
#final_genes


```
Skinny: The reduced data frame with only the relevant genes

This Chunk shows the names of the genes that made it to the reduced data frame
```{r chunky, echo=FALSE}

# Displays names of the final genes.
final_genes = interesting_ens[na.omit(matching)]
final_genes

```



