---
title: "PCA_pathwaytake2"
author: "linh"
date: '2022-06-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "PCAPathwayMatrix"
author: "linh"
date: '2022-06-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load datas}


CompleatePathwayDF = readRDS("C:/Users/trinh/Documents/GitHub/DATATSETS/CompleatePathwayDF.RDS.rds" )

```



```{r PCA for matrix}
# example code from Yoana
#transponseSkinnyBoy=t(SkinnyBoy)
#TRpca=prcomp(transponseSkinnyBoy, retx = TRUE, center = TRUE, scale. = FALSE,
       #tol = NULL, rank. = 14)

# install package
install.packages("ggfortify")
library(ggfortify) 
 
# PCA

Pathway_PCA = prcomp(CompleatePathwayDF_df, retx = TRUE, center = TRUE, scale. = FALSE, tol = NULL, rank. = 2)

CompleatePathwayDF_df = CompleatePathwayDF [1:nrow (CompleatePathwayDF), 1: 9741]


# plot the PCA
library(ggplot2)

ggplot(Pathway_PCA)
autoplot(Pathway_PCA)


## try to transpose it -> each dot is patient

transponsePathwayDF=t(CompleatePathwayDF_df)
Pathway_PCA_patient = prcomp(transponsePathwayDF, retx = TRUE, center = TRUE, scale. = TRUE, tol = NULL, rank. = 2)

autoplot(Pathway_PCA_patient)

### -log10 from p value
pathway_df_Scale = -log10(CompleatePathwayDF_df)

transponsePathwayDF_scale = t(pathway_df_Scale)

Pathway_PCA_patient = prcomp(transponsePathwayDF_scale, retx = TRUE, center = TRUE, scale. = TRUE, tol = NULL, rank. = 10)

install.packages("ggfortify")
library(ggfortify)

autoplot(Pathway_PCA_patient)

library(ggplot2)
ggplot2::ggplot(data = Pathway_PCA_patient, mapping = aes(x = Pathway_PCA_patient$x[,1], y = Pathway_PCA_patient$x[,2])) + geom_point(alpha = 0.2, color = "black")


