---
title: "CompareGenesWithBoxplot"
author: "Ekin Ören"
date: '2022-05-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This chunk finds very highly expressed genes
```{r pressure, echo=FALSE}

#load skinny boy
SkinnyBoy = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/SkinnyBoy.RDS")


#________________________________________________________________________________________________
#Makes a dataframe of highly expressed genes and their expression mean SKİNNYBOY
A = rowMeans(SkinnyBoy) 

RowMeansSkinnyBoy= c()
for (i in 1:14674){  RowMeansSkinnyBoy[i] = A[[i]] } #puts the rowmeans in a vector

quantile(RowMeansSkinnyBoy) #We take the treshold value below from the quantiles
Treshold = 10

Chonk = data.frame()
for (i in 1:14674) {
  
  if(RowMeansSkinnyBoy[i]> Treshold){ 
    
    Chonk[i,1] = names(A)[i]
    Chonk[i,2] = A[[i]]
    
    }
}

FullChonkSB = Chonk[complete.cases(Chonk), ]
dim(FullChonkSB) # 22 genes in treshold = 10
#________________________________________________________________________________________________



#________________________________________________________________________________________________
#Makes a dataframe of highly expressed genes and their expression mean SADTT
A1 = rowMeans(SkinnySadTT) 

RowMeansSkinnySadTT= c()
for (i in 1:14674){  RowMeansSkinnySadTT[i] = A[[i]] } #puts the rowmeans in a vector

quantile(RowMeansSkinnySadTT) #We take the treshold value below from the quantiles
Treshold = 10

Chonk1 = data.frame()
for (i in 1:14674) {
  
  if(RowMeansSkinnySadTT[i]> Treshold){ 
    
    Chonk1[i,1] = names(A)[i]
    Chonk1[i,2] = A[[i]]
    
    }
}

FullChonkSadTT = Chonk1[complete.cases(Chonk1), ]
dim(FullChonkSadTT) # 22 genes in treshold = 10
#________________________________________________________________________________________________



#________________________________________________________________________________________________
#Makes a dataframe of highly expressed genes and their expression mean  HappyTT
A2 = rowMeans(SkinnyHappyTT) 

RowMeansSkinnyHappyTT= c()
for (i in 1:14674){  RowMeansSkinnyHappyTT[i] = A[[i]] } #puts the rowmeans in a vector

quantile(RowMeansSkinnyHappyTT) #We take the treshold value below from the quantiles
Treshold = 10

Chonk2 = data.frame()
for (i in 1:14674) {
  
  if(RowMeansSkinnyBoy[i]> Treshold){ 
    
    Chonk2[i,1] = names(A)[i]
    Chonk2[i,2] = A[[i]]
    
    }
}

FullChonkHappyTT = Chonk2[complete.cases(Chonk2), ]
dim(FullChonkHappyTT) # 22 genes in treshold = 10
head(FullChonkSB)
FullChonkSB[1,]
#________________________________________________________________________________________________


```

Creates a boxplot of the 22 genes that are expressed the highest.
```{r setup, include=FALSE}

boxplot(FullChonkSB[,2], FullChonkSadTT[,2], FullChonkHappyTT[,2],
main = "Comparison of most expressed genes",
at = c(1,2,3),
names = c("SkinnyBoy", "SadTT", "HappyTT"),
las = 2,
col = c("violet","blue", "pink"),
border = "black",
horizontal = TRUE,
notch = TRUE
)


```



This chunk finds very varient genes for our three datasets
```{r pressure, echo=FALSE}

#load skinny boy
SkinnyBoy = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/SkinnyBoy.RDS")


#________________________________________________________________________________________________
#Makes a dataframe of highly Varient genes and their expression mean SKİNNYBOY

varianceA = apply(SkinnyBoy, 1, var)
VeryVariantSB = SkinnyBoy[varianceA > quantile(varianceA , probs = 0.999), ]
dim(VeryVariantSB)

#________________________________________________________________________________________________

#________________________________________________________________________________________________
#Makes a dataframe of highly Varient genes and their expression mean SKİNNYSADTT

varianceB = apply(SkinnySadTT, 1, var)
VeryVariantSkinnySadTT = SkinnySadTT[varianceB > quantile(varianceB , probs = 0.999), ]
dim(VeryVariantSkinnySadTT)

#________________________________________________________________________________________________

#________________________________________________________________________________________________
#Makes a dataframe of highly Varient genes and their expression mean SKİNNYHAPPYTT

varianceC = apply(SkinnyHappyTT, 1, var)
VeryVariantSkinnyHappyTT = SkinnyHappyTT[varianceC > quantile(varianceC , probs = 0.999), ]
dim(VeryVariantSkinnyHappyTT)

#________________________________________________________________________________________________

   
```



Generate mean expression vecotrs
```{r setup, include=FALSE}

Ali1 = rowMeans(VeryVariantSB) 
RowMeansVVSB= c()
for (i in 1:15){  RowMeansVVSB[i] = Ali1[[i]] } #puts the rowmeans in a vector
RowMeansVVSB

Ali2 = rowMeans(VeryVariantSkinnySadTT) 
RowMeansVVSadTT= c()
for (i in 1:19){  RowMeansVVSadTT[i] = Ali2[[i]] } #puts the rowmeans in a vector
RowMeansVVSadTT

Ali3 = rowMeans(VeryVariantSkinnyHappyTT) 
RowMeansVVHappyTT= c()
for (i in 1:19){  RowMeansVVHappyTT[i] = Ali3[[i]] } #puts the rowmeans in a vector
RowMeansVVHappyTT

```







Creates a boxplot of the the genes with the greatest varience. brca tum norm and skinnyboy
```{r setup, include=FALSE}


boxplot(RowMeansVVSB, RowMeansVVSadTT, RowMeansVVHappyTT,
main = "Comparison of mean expressions of most variant genes",
at = c(1,2,3),
names = c("SkinnyBoy", "SadTT", "HappyTT"),
las = 2,
col = c("purple","lightblue", "pink"),
border = "black",
horizontal = TRUE,
notch = FALSE
)

```




This chunk finds very varient genes for all tumour types
```{r pressure, echo=FALSE}

BRCAraw = tcga_tumor_norm$BRCA$tumor
KIRCraw = tcga_tumor_norm$KIRC$tumor
LUADraw = tcga_tumor_norm$LUAD$tumor
PRADraw = tcga_tumor_norm$PRAD$tumor
THCAraw = tcga_tumor_norm$THCA$tumor


#________________________________________________________________________________________________

var1 = apply(BRCAraw, 1, var)
VeryVariantBRCA = BRCAraw[var1 > quantile(var1 , probs = 0.999), ]
dim(VeryVariantBRCA)

#________________________________________________________________________________________________

#________________________________________________________________________________________________

var2 = apply(KIRCraw, 1, var)
VeryVariantKIRC= KIRCraw[var2 > quantile(var2, probs = 0.999), ]
dim(VeryVariantKIRC)

#________________________________________________________________________________________________

#________________________________________________________________________________________________

var3 = apply(LUADraw, 1, var)
VeryVariantLUAD = LUADraw[var3 > quantile(var3 , probs = 0.999), ]
dim(VeryVariantLUAD)

#________________________________________________________________________________________________

#________________________________________________________________________________________________

var4 = apply(PRADraw, 1, var)
VeryVariantPRAD = PRADraw[var4 > quantile(var4 , probs = 0.999), ]
dim(VeryVariantPRAD)

#________________________________________________________________________________________________

#________________________________________________________________________________________________

var5 = apply(THCAraw, 1, var)
VeryVariantTHCA = THCAraw[var5 > quantile(var5 , probs = 0.999), ]
dim(VeryVariantTHCA)
head(VeryVariantTHCA)
#________________________________________________________________________________________________

   
```


Generate mean expression vecotrs for each
```{r setup, include=FALSE}

oo1 = rowMeans(VeryVariantBRCA) 
RowMeans1= c()
for (i in 1:20){  RowMeans1[i] = oo1[[i]] } #puts the rowmeans in a vector
RowMeans1

oo2 = rowMeans(VeryVariantKIRC) 
RowMeans2= c()
for (i in 1:20){  RowMeans2[i] = oo2[[i]] } #puts the rowmeans in a vector
RowMeans2

oo3 = rowMeans(VeryVariantLUAD) 
RowMeans3= c()
for (i in 1:20){  RowMeans3[i] = oo3[[i]] } #puts the rowmeans in a vector
RowMeans3

oo4 = rowMeans(VeryVariantPRAD) 
RowMeans4= c()
for (i in 1:20){  RowMeans4[i] = oo4[[i]] } #puts the rowmeans in a vector
RowMeans4

oo5 = rowMeans(VeryVariantTHCA) 
RowMeans5= c()
for (i in 1:20){  RowMeans5[i] = oo5[[i]] } #puts the rowmeans in a vector
RowMeans5

AnnotVector = c("BRCA","KIRC","LUAD","PRAD","THCA")

```



Creates a boxplot of the the genes with the greatest varience amongst the five cancer types.
```{r setup, include=FALSE}

boxplot(RowMeans1, RowMeans2, RowMeans3, RowMeans4, RowMeans5,
main = "mean expressions of most variant genes amongst cancer types",
at = c(1,2,3,4,5),
names = AnnotVector,
las = 2,
col = c("#7B0D1E", "pink", "violet", "light blue", "#35605A"),
border = "black",
horizontal = TRUE,
notch = FALSE
)

#Same as atripchart below

x <- as.list(VeryVariantTHCA)
ZOOT <- list(RowMeans1, RowMeans2, RowMeans3, RowMeans4, RowMeans5)

stripchart(ZOOT,
main="mean expression of the most varient genes amongst cancer types",
col = c("#7B0D1E", "pink", "violet", "light blue", "#35605A"),
group.names = AnnotVector,
vertical=TRUE,
pch=16
)

```







