---
title: "Yoana's Code"
author: "Ekin Ören"
date: '2022-05-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


First part, plot for variance again mean in only biotype filtered data
```{r pressure, echo=FALSE}

#Loading the data
#LOAD ONLY BIOTYPE CLEANED DATA

sapply.variance<-sapply(BiotypeFilteredBoy, var)
sapply.mean<-sapply(BiotypeFilteredBoy, mean)
dim(BiotypeFilteredBoy)
head(BiotypeFilteredBoy)

```


in order to use ggpots you might need rtools for version 4.2 _ https://cran.r-project.org/bin/windows/Rtools/
Then install  tidyverse package and then ggplot

```{r pressure, echo=FALSE}

#Using ggplot
install.packages("Rcmdr", dependencies = TRUE)
install.packages("tidyverse")
library(ggplot2)

```




First part, plot for variance again mean in bigboy data
```{r pressure, echo=FALSE}

#Loading the data
#LOAD ONLY BIOTYPE CLEANED DATA

sapply.variance0<-apply(BigBoy, 1, var)
sapply.mean0<-apply(BigBoy, 1, mean)


SkinnyBoy = readRDS("C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/SkinnyBoy.RDS")

dim(SkinnyBoy)

```


```{r pressure, echo=FALSE}

Mean<-(sapply.mean0)
Variance<-(sapply.variance0)
mean.and.variance.sapply <- data.frame(Variance , Mean)
myplot<-ggplot(data=mean.and.variance.sapply,aes(x= Mean, y=Variance, geom))+ geom_point()+
  geom_point(data =mean.and.variance.sapply, colour = 'red', size = 2) +ylim(0,50) +xlim(-10, 15)
print(myplot)

```



First part, plot for variance again mean in only biotype filtered data
```{r pressure, echo=FALSE}

#Loading the data
#LOAD ONLY BIOTYPE CLEANED DATA


sapply.variance1<-apply(BiotypeFilteredBoy, 1, var)
sapply.mean1<-apply(BiotypeFilteredBoy, 1, mean)

```


```{r pressure, echo=FALSE}

Mean<-(sapply.mean1)
Variance<-(sapply.variance1)
mean.and.variance.sapply <- data.frame(Variance , Mean)
myplot<-ggplot(data=mean.and.variance.sapply,aes(x= Mean, y=Variance, geom))+ geom_point()+
  geom_point(data =mean.and.variance.sapply, colour = 'red', size = 2) +ylim(0,50) +xlim(-10, 15)
print(myplot)

```

_____________________________________________________________________________________________________________________________________
Doing it All over again for SkinnyBoy




First part, plot for variance again mean in only biotype filtered data
```{r pressure, echo=FALSE}

#Loading the data
#LOAD ONLY BIOTYPE CLEANED DATA

sapply.variance2<-apply(SkinnyBoy, 1, var)
sapply.mean2<-apply(SkinnyBoy, 1, mean)


```


```{r pressure, echo=FALSE}

Mean<-(sapply.mean2)
Variance<-(sapply.variance2)
mean.and.variance.sapply <- data.frame(Variance , Mean)
myplot<-ggplot(data=mean.and.variance.sapply,aes(x= Mean, y=Variance, geom))+ geom_point()+
  geom_point(data =mean.and.variance.sapply, colour = 'red', size = 2)  +ylim(0,50) +xlim(-10, 15)
print(myplot)


#saveRDS(SkinnyBoy, file = "C:/Users/EKN/Documents/GitHub/2022-topic-02-team-01/SkinnyBoy.RDS")
#SkinnyBoy
```




