---
title: "VisualizeExpressionsHistogram"
author: "Ekin Ören"
date: '2022-05-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

WARNİNG: For this code to run, you need to have "BigBoy" dataframe and "tcga_tumor_norm" datalist in your enviroment.if not, go load em.


Calculating the means of each row of the big boy, BRCA and norm-tissue BRCA gene expression data.
in the end we get a vector of the mean expression of each gene in these dataframes. (one vector for each data frame)
```{r Histogrammy, include=FALSE}

SadBoobExpressions = tcga_tumor_norm$BRCA$tumor
HappyBoobExpressions = tcga_tumor_norm$BRCA$normal

ExpressionMeansBigBoy = rowMeans(BigBoy)
ExpressionMeansSadBoob = rowMeans(SadBoobExpressions)
ExpressionMeansHappyBoob = rowMeans(HappyBoobExpressions)

ExpressionMeansBigBoyLog = log(rowMeans(BigBoy))
ExpressionMeansSadBoobLog = log(rowMeans(SadBoobExpressions))
ExpressionMeansHappyBoobLog = log(rowMeans(HappyBoobExpressions))

```


Creating the Histogramms of gene expression for the named datasets. There is also histograms of logs of the same vecotrs
y axis is frequency of expression in this range, and x axis is expression rates.
```{r Histogrammy, include=FALSE}

hist(ExpressionMeansBigBoy, xlab = "", main="Means of Cancer Gene Expression Data", breaks = 100, col = "purple")
hist(ExpressionMeansHappyBoob, xlab = "", main="Means of Breast Cancer Patients Healthy Tissue Gene Expression Data", breaks = 100, col = "pink", ylim = c(0,1500))
hist(ExpressionMeansSadBoob, xlab = "", main="Means of Breast Cancer Tissue Gene Expression Data", breaks = 100, col = "violet", ylim = c(0,1500))


hist(ExpressionMeansBigBoyLog, xlab = "", main="Means of Cancer Gene Expression Data Log", breaks = 100, col = "purple")
hist(ExpressionMeansHappyBoobLog, xlab = "", main="Means of Breast Cancer Patients Healthy Tissue Gene Expression Data Log", breaks = 100, col = "pink", ylim = c(0,1500))
hist(ExpressionMeansSadBoobLog, xlab = "", main="Means of Breast Cancer Tissue Gene Expression Data Log", breaks = 100, col = "violet", ylim = c(0,1500))

```









